---
title: "Rebecca's OB PN finding binary Biomarkers between data iterations"
author: "RebeccaKChance" "Diya Das"
date: "June 13, 2018"
utput:
  html_document:
    fig_height: 7
    fig_width: 7
    toc: yes
    code_folding: hide
    toc_float: yes
editor_options:
  chunk_output_type: console
---
```{r options, include=FALSE}
knitr::opts_chunk$set(cache=TRUE, error=FALSE, message=FALSE, warning=FALSE)

## Update bioconductor and clusterExperiment

setwd("C:/RCOB")
install.packages('installr')
library("installr")
installr()  #remove old Rs control panel programs R so libraries aren't confused
# installing/loading the latest installr package:
install.packages("installr"); library(installr) # install+load installr
 
updateR() # updating R.


library(devtools)
assignInNamespace("version_info", c(devtools:::version_info, list("3.5" = list(version_min = "3.3.0", version_max = "99.99.99", path = "bin"))), "devtools")
find_rtools() # is TRUE now

source("https://bioconductor.org/biocLite.R")
biocLite() #update bioconductor install
BiocManager::install
#biocLite("BiocUpgrade") #really update
biocValid()  
biocLite("pkgmaker")
library(BiocInstaller)
devtools::install_github('renozao/NMF@devel') #dev version NMF

library(devtools)
install.packages("stringi",type="win.binary") #if problems remove old documents>R>3.5>lib
library(stringi)
packageVersion("stringi")
library(MASS, mgcv)
source("https://bioconductor.org/biocLite.R")
BiocManager::install("clusterExperiment", dependencies = TRUE) #2.0.2 June 2018 #2.1.6 Oct 2018
library(clusterExperiment)
packageVersion("clusterExperiment")
```
#Load clusters and colors
```{r readin}
colb <- c("dodgerblue", "purple", "green") #pal <- clusterExperiment::bigPalette
cole <- c("blue","magenta")
colRKC <- c("deepskyblue", "deeppink", "darkviolet", "gold", "chocolate4", "pink", "darkorange2", "chartreuse3", "slategray2", "firebrick3")

############-----------load required data for plots: counts matrix, clusterExperiment object for heatmaps, cluster vector, qc data
library(readr)
load("forGeneExpression_tSNEs.Rda") #save(sconeNorm, ce, vargenes, batch, expt, resolution, tsne_scone, scone50_cl, z_cl, qc, file="forGeneExpression_tSNEs.Rda")
load("RCOB2AB5S_sconetSNE_ce_otherData.Rda")
load("RCOB2AB5S_sconeClusterMany.Rda") # from RSEC May 26 2018(from RCOB2AB5S_cluster.R)  #cl

```

Now we have formalized into automated code what I did manually for two years by screening heatmaps

#Oct4.18 DD made our best version yet:{r eda}
############ pickBinary ####################

```

```{r pickBinary}
cols_all <-apply(transformData(tmp),1, function(x) tapply(x, primaryCluster(tmp), mean))
cols_all <- t(cols_all) #matrix with average exp of genes (row) by cluster (column)

# dim(cols_all) #9631 by 51
cols_all <- cols_all[,2:51] #check cluster number here
# head(cols_all)
# head(de_onevall$ContrastName)
ClusLabels <-levels(de_onevall$ContrastName)
colnames(cols_all) <- ClusLabels
# head(cols_all)
# #head(cols_all[Cl24,])
# m24s<-m24_DE$Feature
# head(cols_all[m24s,])


finalmarker <- lapply(levels(de_onevall$ContrastName), function(contrast){
DEgenes <- de_onevall %>% filter(ContrastName==contrast) %>% 
  filter(logFC > 0) %>%
  arrange(desc(logFC)) %>% 
  select(Feature) %>% unlist()
sub_mat <- cols_all[DEgenes,setdiff(colnames(cols_all), contrast)]
DEgenes <- rownames(sub_mat)[apply(sub_mat, 1, function(x) all(x < 1.8))]
#if (length(DEgenes == 0)) DEgenes <- NA
return(DEgenes)
})
names(finalmarker) <- levels(de_onevall$ContrastName)

head(finalmarker$Cl01[1:50])
```
# Lock it down
```{r save_filtered}
save(finalmarker, file="RCOB2AB5S_finalMarker.rda")
```


#finding clusterID based on gene expression
#for quickly comparing cluster ID between datasets if have a known marker gene

```{r mappingKnownMarkers}
message(paste(grep("Rhoc", finalmarker), collapse = ", ")) 
paste(finalmarker[c(2,4,25)])
message(paste(finalmarker[c(2,4,25)], collapse = ", "))

Cplx3common<-intersect(unlist(finalmarker$Cl14), unlist(finalmarker$Cl35))
Cplx3common<-intersect(Cplx3common, unlist(finalmarker$Cl12))
#not 2,3, 5 have 24 (2 and 3 v. common ~300) #17 good pick also has 11
grep("Cplx3", Cplx3common)

clus_outEPL<- paste0(de_onevall$ContrastName[c(14, 3758, 6491, 10766, 13074, 14250, 16575, 18018, 21616, 21916, 22273, 22626, 25652, 25928, 27360, 27950, 31971, 37677, 39313, 44976, 46008, 47747, 53406, 54788, 62600, 63499, 64100)])

  for (i in 1:length(clus_outEPL[])) {
 clus_DE<-unlist(finalmarker[paste0(clus_outEPL[i])], use.names=FALSE)
message(paste(grep("Cplx3", clus_DE), collapse = ", "))
  }


# cluster averages 
M24_DE_avge <- as.data.frame(cols_all[m24s,])
head(M24_DE_avge)
typeof(M24_DE_avge)
dim(M24_DE_avge)
#newdata <- subset(mydata, sex=="m" & age > 25, select=weight:income)
M24_DE_avge<- M24_DE_avge[order(M24_DE_avge[,24], decreasing=TRUE),]

M24_DE_avge24<-data.frame()
M24_DE_avge24$gene <-M24_DE_avge[,1]
M24_DE_avge24$avge <-M24_DE_avge[,24]
head(M24_DE_avge24)

M24_DE_avge$AvgExp <- (cols_all[,24])
head(M24_DE_avge)
dim(M24_DE_avge)
colnames(M24_DE_avge[,1])<-  "AvgExp"
M24_DE_avge<- M24_DE_avge[order(M24_DE_avge[,24], decreasing=TRUE),]
typeof(M24_DE_avge)
head(M24_DE_avge)
dim(M24_DE_avge)
  
a <- list() #make a list
a$M1 <- m1_DE #add one cluster's DE list via data object character string
a$M39 <- m39_DE
a$M40 <- c("Cgnl1", "Tubb6")

#MOB MCL is Cl05 Cl08 Cl24 Cl45
clus_MOBmcl<- c("Cl05", "Cl08", "Cl24", "Cl45")
```

## then loop a heatmap pdf
```{r loopHeatmap}
for (i in 1:length(clus_outEPL[])) {
 clus_DE<-unlist(finalmarker[paste0(clus_outEPL[i])], use.names=FALSE)
 shortlist_clus_DE<-clus_DE[1:50]
  
 NMF::nmf.options(grid.patch=TRUE)
 pdf(paste0("output/viz/RCOB2AB5S_heatmap_rsec_1vall_",clus_outEPL[i],".pdf"), height=8.5, width=11)
 plotHeatmap(tmp, clusterSamplesData="dendrogramValue",
            whichClusters="primaryCluster", 
            clusterFeaturesData=rownames(tmp)[rownames(tmp) %in% shortlist_clus_DE],  fontsize=8,
            main=paste0("RCOB2AB5S_heatmap_1vall_rsec_shortlist_",clus_outEPL[i],"_DE"), breaks=.99 ,  
            labCol=rep("",ncol(tmp)) )
dev.off()


}

```

# Session Info

```{r}
sessionInfo()
```
