---
title: "Rebecca's Olfactory Bulb Projection Neurons as of the end of February 2018"
author: "Russell"
date: "02/28/2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r import, echo=FALSE, cache=TRUE, warning=FALSE, error=FALSE, message=FALSE}

library(SummarizedExperiment)
library(scater)
library(scone)
library(clusterExperiment)
library(ggplot2)
library(magrittr)
library(cowplot)
library(Rtsne)
library(zinbwave)
library(cellrangerRkit)
library(Hmisc)
library(Seurat)
NMF::nmf.options(grid.patch=TRUE)##put anywhere in the script to remove first (blank) page of heatmap pdfs

exptStr <- "RCOB2AB5"
pal <- clusterExperiment::bigPalette
colb <- c("dodgerblue", "goldenrod", "magenta", "purple", "green")
cole <- c("blue","magenta")
colRKC <- c("deepskyblue", "deeppink", "darkviolet", "gold", "chocolate4", "pink", "darkorange2", "chartreuse3", "slategray2", "firebrick3", "gray", "black", "navy", "lightgoldenrod1", "gray20")
set.seed(9887)

mydata <- load_cellranger_matrix("cellrangerOut/RCOB2A/")
RCOB2A <-as.matrix(exprs(mydata))
mydata <- load_cellranger_matrix("cellrangerOut/RCOB2B/")
RCOB2B <-as.matrix(exprs(mydata))
mydata <- load_cellranger_matrix("cellrangerOut/RCOB3A/")
RCOB3A <-as.matrix(exprs(mydata))
mydata <- load_cellranger_matrix("cellrangerOut/RCOB3B/")
RCOB3B <-as.matrix(exprs(mydata))
mydata <- load_cellranger_matrix("cellrangerOut/RCOB4/")
RCOB4 <-as.matrix(exprs(mydata))

mydata <- load_cellranger_matrix("cellrangerOut/RCOBaggregate2-4/")
RCOB <- as.matrix(exprs(mydata))
# remove1 <- length(colnames(RCOB1))
# RCOB <- RCOB[,-c(1:remove1)]


expt <- as.factor(paste0("OB_", c(rep("Slc17a7", NCOL(RCOB)))))
batch <- as.factor(paste0("OB_", c(rep("2A", NCOL(RCOB2A)), rep("2B", NCOL(RCOB2B)), rep("3A", NCOL(RCOB3A)), rep("3B", NCOL(RCOB3B)), rep("4", NCOL(RCOB4)))))

se <- SummarizedExperiment(list(counts=RCOB), colData=data.frame(batch=batch, expt=expt))
genes <- read.table(file="cellrangerOut/RCOBaggregate2-4/outs/filtered_gene_bc_matrices_mex/GRCm38p4Mm10/genes.tsv")
colnames(genes) <- c("geneID", "Symbol")
rowData(se) <- genes
head(rowData(se))
head(colData(se))
rm(RCOB2A, RCOB2B, RCOB3A, RCOB3B, RCOB4)

```

```{r functions, echo=FALSE}
library(rARPACK)
fastpca <- function(expr, scale=FALSE) {
  k <- 50
  svd_raw <- svds(scale(t(expr), center=TRUE, scale=scale), k=k, nu=k, nv=0)
  pc_raw <- svd_raw$u %*% diag(svd_raw$d[1:k])
  return(pc_raw)
}

norm10x = function (ei)
{
  sums = colSums(ei)
  eo = t(t(ei)*median(sums)/sums)
  return(eo)
}
```

# Exploratory Data Analysis
```{r EDA, echo=FALSE}

hist(colSums(assay(se)), breaks=30, 
     xlab='Number of UMI', main="Number of UMI per sample")

hist(colSums(assay(se)>0)/nrow(se), breaks=30, 
     xlab="Proportion of detected genes", main="Proportion of detected genes")

hist(colSums(assay(se)>0), breaks=50, 
     xlab="Number of detected genes", main="Number of detected genes")

hist(colMeans(assay(se)==0), breaks=30, 
     xlab="Proportion of zeros", main="Proportion of zeros")

boxplot(colSums(assay(se))~colData(se)$batch, main="Number of UMI per sample")
boxplot(colSums(assay(se)>0)/nrow(se)~colData(se)$batch, main="Proportion of detected genes")
boxplot(colMeans(assay(se)==0)~colData(se)$batch, main="Proportion of zeros")

simple <- assay(se)[rowSums(assay(se))>10,]
simple <- SUM_FN(simple)

runQCmetrics <- TRUE
if (runQCmetrics) {
pca <- prcomp(t(simple), scale. = TRUE)
screeplot(pca, type = "lines", npcs = 50, main = "Expression PCA screeplot")
plot(pca$x, col=alpha(colb[batch], 0.3), pch=19, cex=0.4, main="simple filtering, scaling --> PCA")
legend("topleft", legend=levels(batch), fill=colb, cex=0.6)

print(paste0("Percent total variance captured in first 50 PCs is ", cumsum(pca$sdev^2/sum(pca$sdev^2))[50] * 100))
  
###---Use scater package to calculate some quality control metrics. Assess QC metrics.

###only works with unique identifiers - Ensembl ids, not gene names b/c some gene name duplication (65/28000)
sce <- newSCESet(countData = assay(se))
sce <- calculateQCMetrics(sce)

rownames(se) <- rowData(se)$Symbol
head(rownames(se))

###only works with gene names
ribo_idx <- grep("^Rpl|^Rps", rowData(se)[,2])
mito_idx <- grep("^Mt", rowData(se)[,2])
ribo_pct <- colSums(assay(se)[ribo_idx,])/colSums(assay(se)) * 100
mito_pct <- colSums(assay(se)[mito_idx,])/colSums(assay(se)) * 100
plot(mito_pct, col=colb[batch], xlab="cell index", main="% mito (Mt*) genes"); legend("topleft", legend=levels(batch), fill=colb, cex=0.8)
boxplot(mito_pct ~ colData(se)$batch, main="percent mito genes", col=colb)
plot(ribo_pct, col=colb[batch], xlab="cell index", main="% ribo (Rpl*) genes"); legend("topleft", legend=levels(batch), fill=colb, cex=0.8)
boxplot(ribo_pct ~ colData(se)$batch, main="percent ribo genes", col=colb)

qc <- as.matrix(data.frame(colData(sce)[,c(2, 4:8)], mito_pct = mito_pct, ribo_pct = ribo_pct))

qcpca <- prcomp(qc, scale. = TRUE)
screeplot(qcpca, type = "lines", main = "QC-PCA screeplot")
plot(qcpca$x, col=alpha(colb[batch], 0.3), pch=19, main="QC PCA")
legend("bottomright", legend=levels(batch), fill=colb, cex=0.8)
# plot(qcpca$x, col=alpha(colb[batch], 0.1), pch=19, main="QC PCA")
# legend("bottomright", legend=levels(batch), fill=colb, cex=0.6)

save(se, sce, qc, pca, qcpca, file="dataObjects/RCOB_unfiltered.Rda")
} else {
  load("dataObjects/RCOB_unfiltered.Rda")
}

fig_data <- data.frame(pca$x[,1:2], qc,
                   QPC1=qcpca$x[,1], QPC2=qcpca$x[,2])

fig_pca <- ggplot(fig_data, aes(x = PC1, y = PC2, color = log10_total_counts)) +
  geom_point() + scale_color_continuous(low = "blue", high = "yellow")
fig_pca

fig_qpca <- ggplot(fig_data, aes(x = QPC1, y = QPC2, color = log10_total_counts)) +
  geom_point() + scale_color_continuous(low = "blue", high = "yellow")
fig_qpca

fig_pca <- ggplot(fig_data, aes(x = PC1, y = PC2, color = ribo_pct)) +
  geom_point() + scale_color_continuous(low = "blue", high = "yellow")
fig_pca

fig_qpca <- ggplot(fig_data, aes(x = QPC1, y = QPC2, color = ribo_pct)) +
  geom_point() + scale_color_continuous(low = "blue", high = "yellow")
fig_qpca

fig_pca <- ggplot(fig_data, aes(x = PC1, y = PC2, color = mito_pct)) +
  geom_point() + scale_color_continuous(low = "blue", high = "yellow")
fig_pca

fig_qpca <- ggplot(fig_data, aes(x = QPC1, y = QPC2, color = mito_pct)) +
  geom_point() + scale_color_continuous(low = "blue", high = "yellow")
fig_qpca

fig_pca <- ggplot(fig_data, aes(x = PC1, y = PC2, color = log10_total_features)) +
  geom_point() + scale_color_continuous(low = "blue", high = "yellow")
fig_pca

fig_qpca <- ggplot(fig_data, aes(x = QPC1, y = QPC2, color = log10_total_features)) +
  geom_point() + scale_color_continuous(low = "blue", high = "yellow")
fig_qpca

rm(mito_idx, ribo_idx, fig_pca, fig_qpca)
```

```{r filtering, echo=FALSE}

colData(se) <- cbind(colData(se), qc)

# select expressed genes only
dim(se)
se <- se[rowSums(assay(se))>0,]
dim(se)

# select common genes
num_reads <- quantile(assay(se)[assay(se) > 0])[4]
num_cells = 0.25*ncol(se)
is_common = rowSums(assay(se) >= num_reads) >= num_cells
table(is_common)

#hk <- read.table("ref/HK_genes.txt")
data("housekeeping")
hk <- capitalize(tolower(housekeeping$V1))
#hk <- read.table("ref/hkl615.txt")
#hk <- as.character(hk[,1])
# hk <- sample(hk, 1000)
# write.table(hk, file="ref/HK_genes_1000sampled.txt", quote = FALSE, col.names = FALSE, row.names = FALSE)
hk <- intersect(hk, rowData(se)$Symbol)
hk_idx <- which(rowData(se)$Symbol %in% hk)

mfilt <- metric_sample_filter(assay(se),
                             nreads = colData(sce)$total_counts,
                             gene_filter = is_common,
                             pos_controls = hk_idx,
                             hard_nreads = 2000,
                             zcut = 3, mixture = FALSE,
                             plot = TRUE)

plot(pca$x, pch=19, col=pal[as.numeric(mfilt$filtered_breadth)+1],
     main = "PCA Filtered on transcriptome 'breadth'")
plot(pca$x, pch=19, col=pal[as.numeric(mfilt$filtered_fnr)+1],
     main = "PCA Filtered on FNR AUC")
plot(pca$x, pch=19, col=pal[as.numeric(mfilt$filtered_nreads)+1],
     main = "PCA Filtered on nreads")

plot(qcpca$x, pch=19, col=pal[as.numeric(mfilt$filtered_breadth)+1],
     main = "QPCA Filtered on transcriptome 'breadth'")
plot(qcpca$x, pch=19, col=pal[as.numeric(mfilt$filtered_fnr)+1],
     main = "QPCA Filtered on FNR AUC")
plot(qcpca$x, pch=19, col=pal[as.numeric(mfilt$filtered_nreads)+1],
     main = "QPCA Filtered on nreads")

table(mfilt$filtered_nreads, mfilt$filtered_fnr)
filter_cell <- !apply(simplify2array(mfilt[!is.na(mfilt)]),1,any)

plot(qcpca$x, pch=19, col=pal[as.numeric(filter_cell)+1],
     main = "PCA Filtered on FNR AUC")


# Final Gene Filtering: Highly expressed in at least 5 cells
num_reads <- quantile(assay(se)[assay(se) > 0])[4]
num_cells = 5
is_quality = rowSums(assay(se) >= num_reads ) >= num_cells
table(is_quality)
filtered <- se[is_quality, filter_cell]
#filtered <- se[is_quality,]
dim(filtered)
rownames(filtered) <- rowData(filtered)$Symbol
qc <- qc[colnames(filtered),]
batch <- colData(filtered)[,1]
names(batch) <- colnames(filtered)
expt <- colData(filtered)[,2]
names(expt) <- colnames(filtered)

```


```{r save_filtered}
save(qc, filtered, batch, expt, file="dataObjects/RCOB_filtered.rda")
```


# Check for batch effects prior to normalization
```{r batch, echo=FALSE}
# check for batch effects
plot(pca$x, pch=19, col=colb[colData(se)$batch],
     main = "PCA Color-coded by batch")
legend("topleft", legend=levels(batch), fill=colb, cex=0.8)
plot(qcpca$x, pch=19, col=colb[colData(se)$batch],
     main = "QPCA Color-coded by batch")
legend("topleft", legend=levels(batch), fill=colb, cex=0.8)

boxplot(pca$x[,1] ~ colData(se)$batch, col=colb)
boxplot(pca$x[,2] ~ colData(se)$batch, col=colb)

simple <- assay(filtered)
#simple <- SUM_FN(simple)

pca <- fastpca(log(simple + 0.1))
tsne_data <- Rtsne(pca[,1:50], pca=FALSE, max_iter = 5000)
plot(tsne_data$Y, pch=19, cex=0.5, col=alpha(colb[colData(filtered)$batch],0.5), main = "Gene filtered, cell filtered, Pre-Norm\n OB cells")
legend("bottomright", legend=levels(batch), fill=colb, cex=0.8)
pdf(file="output/viz/filteredPreNorm/tSNE_batch_postFilt_preNorm.pdf")
plot(tsne_data$Y, pch=19, cex=0.5, col=alpha(colb[colData(filtered)$batch],0.5), main = "Gene filtered, cell filtered, Pre-Norm\n OB cells")
legend("bottomright", legend=levels(batch), fill=colb, cex=0.8)
dev.off()
plot(pca[,1:2], pch=19, col=colb[colData(filtered)$batch])
legend("bottomright", legend=levels(batch), fill=colb, cex=0.8)
pdf(file="output/viz/filteredPreNorm/PCAcoloredByBatch")
plot(pca[,1:2], pch=19, col=colb[colData(filtered)$batch])
legend("bottomright", legend=levels(batch), fill=colb, cex=0.8)
dev.off()


```

# Gene expression plots post filtering, but pre-normalization
```{r preNorm_postFiltering_Plots, echo=FALSE}

filtCounts <- assay(filtered)
logfiltCounts <- log2(filtCounts+1)

tsne_data <- Rtsne(pca[,1:50], pca=FALSE, max_iter = 5000)
plot(tsne_data$Y, pch=19, cex=0.5, col=alpha(colb[colData(filtered)$batch],0.5), main = "Gene filtered, cell filtered, Pre-Norm\n OB cells")
legend("bottomright", legend=levels(batch), fill=colb, cex=0.8)
fig_data <- data.frame(TSNE1=tsne_data$Y[,1], TSNE2=tsne_data$Y[,2],
                   Slc17a7 = logfiltCounts[which(rowData(filtered)$Symbol == "Slc17a7"),],
                   Gad1 = logfiltCounts[which(rowData(filtered)$Symbol == "Gad1"),],
                   tdTomato = logfiltCounts[which(rowData(filtered)$Symbol == "tdTomato"),],
                   Gad2 = logfiltCounts[which(rowData(filtered)$Symbol == "Gad2"),],
                   Fxyd7 = logfiltCounts[which(rowData(filtered)$Symbol == "Fxyd7"),],
                   Reln = logfiltCounts[which(rowData(filtered)$Symbol == "Reln"),],
                   Coch = logfiltCounts[which(rowData(filtered)$Symbol == "Coch"),],
                   Fxyd6 = logfiltCounts[which(rowData(filtered)$Symbol == "Fxyd6"),])


ggplot(fig_data, aes(x = TSNE1, y = TSNE2, color = Slc17a7)) +
  geom_point() + scale_color_continuous(low = "blue", high = "yellow")
ggplot(fig_data, aes(x = TSNE1, y = TSNE2, color = Gad1)) +
  geom_point() + scale_color_continuous(low = "blue", high = "yellow")
ggplot(fig_data, aes(x = TSNE1, y = TSNE2, color = Gad2)) +
  geom_point() + scale_color_continuous(low = "blue", high = "yellow")
ggplot(fig_data, aes(x = TSNE1, y = TSNE2, color = tdTomato)) +
  geom_point() + scale_color_continuous(low = "blue", high = "yellow")
ggplot(fig_data, aes(x = TSNE1, y = TSNE2, color = Reln)) +
  geom_point() + scale_color_continuous(low = "blue", high = "yellow")
ggplot(fig_data, aes(x = TSNE1, y = TSNE2, color = Fxyd6)) +
  geom_point() + scale_color_continuous(low = "blue", high = "yellow")
ggplot(fig_data, aes(x = TSNE1, y = TSNE2, color = Fxyd7)) +
  geom_point() + scale_color_continuous(low = "blue", high = "yellow")
ggplot(fig_data, aes(x = TSNE1, y = TSNE2, color = Coch)) +
  geom_point() + scale_color_continuous(low = "blue", high = "yellow")


OBgenes <- read.table(file="ref/OBgenes.txt", header = TRUE)
poscon <- unique(intersect(OBgenes$Gene, rowData(filtered)$Symbol))
rowData(filtered)$poscon <- (rowData(filtered)$Symbol %in% poscon)

negconeval <- sample(hk, length(poscon))
negconruv <- setdiff(hk, negconeval)

rowData(filtered)$negcon_eval <- (rowData(filtered)$Symbol %in% negconeval)
rowData(filtered)$negcon_ruv <- (rowData(filtered)$Symbol %in% negconruv)

save(filtered, batch, expt, hk, negconeval, negconruv, poscon, file="dataObjects/RCOB_filtered_preScone.Rda")

pdf(file="output/viz/filteredPreNorm/preNorm_poscon_heatmap.pdf", title = "poscon heatmap")
plotHeatmap(logfiltCounts[rowData(filtered)$poscon,], sampleData=data.frame(batch = colData(filtered)[,1], expt = colData(filtered)[,2] ), clusterLegend = list(batch=colb, expt=cole), main="Positive controls", breaks=.99)
dev.off()
plotHeatmap(logfiltCounts[rowData(filtered)$poscon,], sampleData=data.frame(batch = colData(filtered)[,1], expt = colData(filtered)[,2]), clusterLegend = list(batch=colb, expt=cole), main="Positive controls", breaks=.99)


pdf(file="output/viz/filteredPreNorm/preNorm_negconeval_heatmap.pdf", title = "negconeval heatmap")
plotHeatmap(logfiltCounts[rowData(filtered)$negcon_eval,], sampleData=data.frame(batch = colData(filtered)[,1], expt = colData(filtered)[,2]), clusterLegend = list(batch=colb, expt=cole), main="Negative controls, for evalution", breaks=.99)
dev.off()
plotHeatmap(logfiltCounts[rowData(filtered)$negcon_eval,],sampleData=data.frame(batch = colData(filtered)[,1], expt = colData(filtered)[,2]), clusterLegend = list(batch=colb, expt=cole),  main="Negative controls, for evaluation", breaks=.99)

pdf(file="output/viz/filteredPreNorm/preNorm_negconruv_heatmap.pdf", title = "negconruv heatmap")
plotHeatmap(logfiltCounts[rowData(filtered)$negcon_ruv,], sampleData=data.frame(batch = colData(filtered)[,1], expt = colData(filtered)[,2]), clusterLegend = list(batch=colb, expt=cole), main="Negative controls, for RUV", breaks=.99)
dev.off()
plotHeatmap(logfiltCounts[rowData(filtered)$negcon_ruv,], sampleData=data.frame(batch = colData(filtered)[,1], expt = colData(filtered)[,2]), clusterLegend = list(batch=colb, expt=cole), main="Negative controls, for RUV", breaks=.99)

```

#####Correlation of QC metrics with gene expression PCs -- preNormalization
```{r prepForNorm, echo=FALSE}
filtPCA <- fastpca(logfiltCounts)
cors <- lapply(1:5, function(i) abs(cor(filtPCA[,i], qc, method="spearman")))
cors <- unlist(cors)
bars <- data.frame(AbsoluteCorrelation=cors,
                   QC=factor(rep(colnames(qc), 5), levels=colnames(qc)),
                   Dimension=as.factor(rep(paste0("PC", 1:5), each=ncol(qc))))

bars %>%
  ggplot(aes(Dimension, AbsoluteCorrelation, group=QC, fill=QC)) +
  geom_bar(stat="identity", position='dodge') +
  scale_fill_manual(values=bigPalette) + ylim(0, 1) + 
  ggtitle("Correlation between QC and expression PCA") -> fig_barplot
fig_barplot
pdf(file="output/viz/filteredPreNorm/corQCandExprPCA.pdf")
fig_barplot
dev.off()

```

# Normalization
### SCONE --> set-up to run on the cluster
```{r scone, eval=FALSE}

run_scone <- TRUE
if(run_scone) {
scone_obj <- SconeExperiment(filtered,
               which_qc = which(colnames(colData(filtered)) %in% colnames(qc)),
               which_negconruv = 5L,
               which_negconeval = 4L,
               which_poscon = 3L,
               which_batch = 1L,
               which_bio = 2L)
save(scone_obj, file="dataObjects/RCOB_scone_obj1.Rda")
} else { 
  load("RCOB_scone_normMats.Rda")
}
```


### ZINB-WaVE set-up
```{r zinbwavePrep, echo = FALSE, eval=TRUE}


vars <- rowVars(logfiltCounts)
names(vars) <- rownames(logfiltCounts)
vars <- sort(vars, decreasing = TRUE)
vargenes <- names(vars)[1:1000]

save(vargenes, filtered, file=paste0("dataObjects/", exptStr, "_forZinbwaveRun.Rda"))
```


```{r prepForSconeOnCluster, eval = FALSE}
eval_scone <- TRUE
if(eval_scone) {
load(paste0(dataDir, exptStr, "_scone_eval_mp.Rda"))
head(scone_obj2@scone_scores)
scores <- scone_obj2@scone_scores
scone_obj2@scone_params <- scone_obj2@scone_params[-c(7:56),] #none,fq,ruv_k=2,no_bio,batch; none,sum,ruv_k=3,no_bio,batch; none,fq,ruv_k=1,no_bio,batch; none,fq,qc_k=3,no_bio,batch; none,sum,qc_k=3,no_bio,batch;  none,fq,ruv_k=2,no_bio,no_batch

scone_obj3 <- scone_obj2 
save(scone_obj3, file=paste0(dataDir, exptStr, "_scone_obj3.Rda"))
}  

# set.seed(1999)
# scone_obj4 <- scone(scone_obj3,
#                     scaling = list(none=identity,
#                                    fq = FQT_FN, sum = SUM_FN),
#                     zero = "postadjust", k_ruv = 3, k_qc=3,
#                     adjust_batch = "yes", adjust_bio = "no", eval_kclust =  3:14, return_norm = "in_memory")

#####-----running scone on selected normalizations (first 6) on the cluster and retrieving it in memory
#####
```







#Visualizing the output of SCONE and ZINB-WaVE
##This is the beginning of the second stage of analysis
```{r setUpForSelectedOBdata, echo=FALSE}
set.seed(9887)
# NMF::nmf.options(grid.patch=TRUE)##put anywhere in the script to remove first (blank) page of heatmap pdfs
library(rARPACK)
fastpca <- function(expr, scale=FALSE) {
  k <- 50
  svd_raw <- svds(scale(t(expr), center=TRUE, scale=scale), k=k, nu=k, nv=0)
  pc_raw <- svd_raw$u %*% diag(svd_raw$d[1:k])
  return(pc_raw)
}

colRKC <- c("deepskyblue", "deeppink", "darkviolet", "gold", "chocolate4", "pink", "darkorange2", "chartreuse3", "slategray2", "firebrick3", "gray", "black", "navy", "lightgoldenrod1", "gray20")
dataDir <- "~/Research/Projects/SCRNASeq/10XG/olfactBulb/dataObjects/"
outDir <- "~/Research/Projects/SCRNASeq/10XG/olfactBulb/output/"
exptStr <- "RCOB2AB5S"
pal <- clusterExperiment::bigPalette
colb <- c("dodgerblue", "goldenrod", "magenta") #, "purple", "green")
cole <- c("blue","magenta")

library(Hmisc)
data("housekeeping")
hk <- capitalize(tolower(housekeeping$V1))
hk <- intersect(hk, rowData(filtered)$Symbol)
OBgenes <- read.table(file="ref/OBgenes.txt", header = TRUE)
poscon <- unique(intersect(OBgenes$Gene, rowData(filtered)$Symbol))
obClusterGenes <- read.table(file="ref/obClusterGenes.txt", header = FALSE)
obClusterGenes <- unique(intersect(obClusterGenes$V1, rowData(filtered)$Symbol))
negconeval <- sample(hk, length(poscon))
negconruv <- setdiff(hk, negconeval)
```



```{r loadingSconeNorm}
load(paste0(dataDir, exptStr, "_filtered.rda"))
filtCounts <- assay(filtered)
load(paste0(dataDir, exptStr, "_scone_normMats.Rda"))
batch <- colData(filtered)[,1]
expt <- colData(filtered)[,2]

sconeNorm <- get_normalized(scone_obj4, "none,sum,ruv_k=3,no_bio,batch")
rownames(sconeNorm) <- rowData(filtered)[,2]
logSconeNorm <- log2(sconeNorm+1)
save(sconeNorm, logSconeNorm, file=paste0(dataDir, exptStr, "_normCounts1.Rda"))
scone1 <- "none,sum,ruv_k=3,no_bio,batch"
#######
sconeNorm2 <- get_normalized(scone_obj4, "none,fq,ruv_k=2,no_bio,batch")
rownames(sconeNorm2) <- rowData(filtered)[,2]
logSconeNorm2 <- log2(sconeNorm2+1)
save(sconeNorm2, logSconeNorm2, file="dataObjects/RCOB_normCounts2.Rda")
scone2 <- "none,fq,ruv_k=2,no_bio,batch"
# #######
# sconeNorm3 <- get_normalized(scone_obj4, "none,fq,ruv_k=3,no_bio,batch")
# rownames(sconeNorm3) <- rowData(filtered)[,2]
# logSconeNorm3 <- log2(sconeNorm3+1)
# save(sconeNorm3, logSconeNorm3, file="dataObjects/RCOB_normCounts3.Rda")

# #######
# sconeNorm4 <- get_normalized(scone_obj4, "none,fq,ruv_k=3,no_bio,no_batch")
# rownames(sconeNorm4) <- rowData(filtered)[,2]
# logSconeNorm4 <- log2(sconeNorm4+1)
# save(sconeNorm4, logSconeNorm4, file="dataObjects/RCOB_normCounts4.Rda")

```


##Correlation of QC metrics with expression PCs for scone norm 1
#####using none,sum,ruv_k=3,no_bio,batch
```{r SCONEnorm_qc, echo=FALSE, eval=TRUE}

# pca_sconeNorm <- prcomp(t(logSconeNorm))
pca_sconeNorm <- fastpca(logSconeNorm)

cors <- lapply(1:5, function(i) abs(cor(pca_sconeNorm[,i], qc, method="spearman")))
cors <- unlist(cors)
bars <- data.frame(AbsoluteCorrelation=cors,
                   QC=factor(rep(colnames(qc), 5), levels=colnames(qc)),
                   Dimension=as.factor(rep(paste0("PC", 1:5), each=ncol(qc))))

bars %>%
  ggplot(aes(Dimension, AbsoluteCorrelation, group=QC, fill=QC)) +
  geom_bar(stat="identity", position='dodge') +
  scale_fill_manual(values=bigPalette) + ylim(0, 1) + 
  ggtitle(paste0(scone1, "\nCorrelation between QC and expression PCA")) -> fig_barplot
fig_barplot
pdf(file=paste0(outDir, "viz/", exptStr, "_", scone1, "_corQCandExprPCA.pdf"))
fig_barplot
dev.off()
################

```

##Correlation of QC metrics with expression PCs for scone norm 2
#####using none,sum,qc_k=3,no_bio,batch
```{r moreSconeNormQC, echo=FALSE, eval=TRUE}
             
###################

# pca_sconeNorm2 <- prcomp(t(logSconeNorm2))
pca_sconeNorm2 <- fastpca(logSconeNorm2)

cors <- lapply(1:5, function(i) abs(cor(pca_sconeNorm2[,i], qc, method="spearman")))
cors <- unlist(cors)
bars <- data.frame(AbsoluteCorrelation=cors,
                   QC=factor(rep(colnames(qc), 5), levels=colnames(qc)),
                   Dimension=as.factor(rep(paste0("PC", 1:5), each=ncol(qc))))

bars %>%
  ggplot(aes(Dimension, AbsoluteCorrelation, group=QC, fill=QC)) +
  geom_bar(stat="identity", position='dodge') +
  scale_fill_manual(values=bigPalette) + ylim(0, 1) + 
  ggtitle(paste0(scone2, "\nCorrelation between QC and expression PCA")) -> fig_barplot
fig_barplot
pdf(file=paste0(outDir, "viz/", exptStr, "_", scone2, "_corQCandExprPCA.pdf"))
fig_barplot
dev.off()

```


## Plots of scone normalized data
####Decided to move forward with the : 
```{r sconeNorm_plot, echo=FALSE, eval=TRUE}
####Decided to go with logSconeNorm2 using qc factors

vars <- rowVars(logSconeNorm)
names(vars) <- rownames(logSconeNorm)
vars <- sort(vars, decreasing = TRUE)
vargenes <- names(vars)[1:1000]

# pca_sconeNorm <- prcomp(t(logSconeNorm[vargenes, ]))
pca_sconeNorm <- fastpca(logSconeNorm[vargenes, ])

tsne_data <- Rtsne(pca_sconeNorm[,1:10], pca=FALSE, max_iter = 5000)

plot(pca_sconeNorm[,1:2], pch=19, col=colb[colData(filtered)$batch])
legend("topleft", legend=levels(batch), fill=colb, cex=0.5)
plot(tsne_data$Y, pch=19, cex = 0.5, col=colb[colData(filtered)$batch], main = "t-SNE: 1Kvar genes, sconeNorm, 10 PCs")
legend("topleft", legend=levels(batch), fill=colb, cex=0.5)

fig_data <- data.frame(PC1=pca_sconeNorm[,1], PC2=pca_sconeNorm[,2], qc, TSNE1=tsne_data$Y[,1], TSNE2=tsne_data$Y[,2],
                   Slc17a7 = logSconeNorm[which(rowData(filtered)$Symbol == "Slc17a7"),],
                   Gad1 = logSconeNorm[which(rowData(filtered)$Symbol == "Gad1"),],
                   tdTomato = logSconeNorm[which(rowData(filtered)$Symbol == "tdTomato"),],
                   Ebf1 = logSconeNorm[which(rowData(filtered)$Symbol == "Ebf1"),],
                   Reln = logSconeNorm[which(rowData(filtered)$Symbol == "Reln"),],
                   Thy1 = logSconeNorm[which(rowData(filtered)$Symbol == "Thy1"),],
                   Coch = logSconeNorm[which(rowData(filtered)$Symbol == "Coch"),])


fig_pca <- ggplot(fig_data, aes(x = PC1, y = PC2, color = log10_total_counts)) +
  geom_point() + scale_color_continuous(low = "blue", high = "yellow")
fig_pca

fig_tsne <- ggplot(fig_data, aes(x = TSNE1, y = TSNE2, color = log10_total_counts)) +
  geom_point() + scale_color_continuous(low = "blue", high = "yellow")
fig_tsne

ggplot(fig_data, aes(x = TSNE1, y = TSNE2, color = Slc17a7)) +
  geom_point() + scale_color_continuous(low = "blue", high = "yellow")
ggplot(fig_data, aes(x = TSNE1, y = TSNE2, color = Gad1)) +
  geom_point() + scale_color_continuous(low = "blue", high = "yellow")
ggplot(fig_data, aes(x = TSNE1, y = TSNE2, color = Ebf1)) +
  geom_point() + scale_color_continuous(low = "blue", high = "yellow")
ggplot(fig_data, aes(x = TSNE1, y = TSNE2, color = tdTomato)) +
  geom_point() + scale_color_continuous(low = "blue", high = "yellow")
ggplot(fig_data, aes(x = TSNE1, y = TSNE2, color = Reln)) +
  geom_point() + scale_color_continuous(low = "blue", high = "yellow")
ggplot(fig_data, aes(x = TSNE1, y = TSNE2, color = Thy1)) +
  geom_point() + scale_color_continuous(low = "blue", high = "yellow")
ggplot(fig_data, aes(x = TSNE1, y = TSNE2, color = Coch)) +
  geom_point() + scale_color_continuous(low = "blue", high = "yellow")

pdf(file=paste0(outDir, "viz/", exptStr, "_logSconeNorm_50VarGenes_heatmap.pdf"), title = "post-SCONEnorm 50 varGene heatmap")
clusterExperiment::plotHeatmap(logSconeNorm[vargenes[1:50],], sampleData=data.frame(batch = colData(filtered)[,1]), clusterLegend = list(batch=colb), main="50 most variable genes", breaks=0.99)
dev.off()
clusterExperiment::plotHeatmap(logSconeNorm[vargenes[1:50],], sampleData=data.frame(batch = colData(filtered)[,1]), clusterLegend = list(batch=colb), main="Scone normalized data, \n50 most variable genes", breaks=.99)

pdf(file=paste0(outDir, "viz/", exptStr, "_logSconeNorm_poscon_heatmap.pdf"), title = paste0(exptStr, " post-SCONEnorm poscon heatmap"))
clusterExperiment::plotHeatmap(logSconeNorm[poscon,], sampleData=data.frame(batch = batch), clusterLegend = list(batch=colb), main="Positive controls", breaks=0.99)
dev.off()
clusterExperiment::plotHeatmap(logSconeNorm[rowData(filtered)$poscon,], sampleData=data.frame(batch = colData(filtered)[,1]), clusterLegend = list(batch=colb), main=paste0(exptStr, " scone normalized data, positive controls"), breaks=.99)
pdf(file=paste0(outDir, "viz/", exptStr, "_logSconeNorm_poscon2_heatmap.pdf"), title = paste0(exptStr, " post-SCONEnorm poscon2 heatmap"))
clusterExperiment::plotHeatmap(logSconeNorm[obClusterGenes,], sampleData=data.frame(batch = batch), clusterLegend = list(batch=colb), main="Positive controls 2", breaks=0.99)
dev.off()
clusterExperiment::plotHeatmap(logSconeNorm[obClusterGenes,], sampleData=data.frame(batch = colData(filtered)[,1]), clusterLegend = list(batch=colb), main=paste0(exptStr, " scone normalized data, positive controls 2"), breaks=.99)

pdf(file=paste0(outDir, "viz/", exptStr, "_logSconeNorm_negconeval_heatmap.pdf"), title = "post-SCONEnorm negative controls heatmap")
clusterExperiment::plotHeatmap(logSconeNorm[rowData(filtered)$negcon_eval,], sampleData=data.frame(batch = colData(filtered)$batch), clusterLegend = list(batch=colb), main="example negative controls", breaks=0.99)
dev.off() 
clusterExperiment::plotHeatmap(logSconeNorm[rowData(filtered)$negcon_eval,],  sampleData=data.frame(batch = colData(filtered)[,1]), clusterLegend = list(batch=colb), main=paste0(exptStr, " scone normalized data, example negative controls"), breaks=.99)
############################

```

#### ZINB-WaVE --> running ZINB-WaVE on the cluster 
```{r zinb, echo = FALSE, eval=FALSE}

run_zinbwave <- TRUE
if(run_zinbwave) {
library(BiocParallel)
library(doParallel)
registerDoParallel(3)
register(DoparParam())


zinb <- zinbFit(filtered[vargenes,], X = "~ log10_total_counts + pct_counts_in_top_200_features + ribo_pct", K=10, epsilon=1000)
save(zinb, file="RCOB2AB5_zinbwave_TC_ribo_batch.rda")
} else {
  load("RCOB_zinbwave_totalCounts.rda")
}
#+ pct_counts_top_500_features
W <- getW(zinb)
d <- dist(W)
tsne_zinb <- Rtsne(d, is_distance = TRUE, pca = FALSE, max_iter=5000)

save(W, file = "RCOB_1Kvar_zinbW.Rda")

```

##t-SNE by batch to visualize the ZINB-WaVE results
```{r visualizingZINBoutput}

library(zinbwave)

importFiltered <- FALSE
if (importFiltered){ 
load(paste0(dataDir, exptStr, "_filtered.rda"))
load(paste0(dataDir, exptStr, "_filteredCounts.rda"))
} 

load(paste0("dataObjects/", exptStr, "_zinbwave1Kvar_TC_top200_batch.Rda"))
load(paste0("dataObjects/", exptStr, "_TC_top200_batch_zinbW.Rda"))
load(paste0("dataObjects/", exptStr, "_TC_top200_batch_tsne_zinb.Rda"))

pdf(file=paste0(outDir, "viz/", exptStr, "_tSNE_ZINBWaVE_byBatch.pdf"), title = "t-SNE of ZINB-WaVE")
plot(tsne_zinb$Y, pch=19, cex=0.4, col=colb[batch], main = "t-SNE (ZINB-WaVE), 1KvarGenes,K=10")
legend("topleft", legend=levels(batch), fill=colb, cex=0.5)
dev.off()
plot(tsne_zinb$Y, pch=19, cex=0.4, col=colb[batch], main = "t-SNE (ZINB-WaVE), 1KvarGenes, K=10")
legend("topleft", legend=levels(batch), fill=colb, cex=0.5)

zinbObj <- zinbwave(filtered[vargenes[1:1000],], fitted_model = zinb, K = 10, epsilon = 1000)

```


###Correlation of QC metrics with ZINB-WaVE W projection components
```{r correlationQC&ZinB}
zbrd <- getW(zinb)
cors <- lapply(1:5, function(i) abs(cor(zbrd[,i], qc, method="spearman")))
cors <- unlist(cors)
bars <- data.frame(AbsoluteCorrelation=cors,
                   QC=factor(rep(colnames(qc), 5), levels=colnames(qc)),
                   Dimension=as.factor(rep(paste0("zinbW", 1:5), each=ncol(qc))))

bars %>%
  ggplot(aes(Dimension, AbsoluteCorrelation, group=QC, fill=QC)) +
  geom_bar(stat="identity", position='dodge') +
  scale_fill_manual(values=bigPalette) + ylim(0, 1) + 
  ggtitle("Correlation between QC and ZINBWaVE W (1Kvar)") -> fig_barplot
fig_barplot
pdf(file=paste0(outDir, "viz/", exptStr, "_zinb1KTCtop200Batch_corQCandZinbW.pdf")) 
fig_barplot
dev.off()

```

## Clustering with SNN on 1. ZINB-WaVE (top 1K variable genes) output; 2. top 50 PCs of 1K most variable genes, scone norm

###SNN clustering on ZINB-WaVE output with top 1K most variable genes
```{r seurat_zinb, echo=FALSE, eval=TRUE}

#colRKC <- c("deepskyblue", "deeppink", "darkviolet", "gold", "chocolate4", "pink", "darkorange2", "chartreuse3", "slategray2", "firebrick3", "magenta")

library(dplyr)
library(Matrix)

#zinbObj <- zinbwave(filtered[vargenes[1:1000],], fitted_model = zinb, K = 10, epsilon = 1000)

seu <- CreateSeuratObject(raw.data = sconeNorm, min.cells = 1, min.genes = 1, 
     project = "RCOB")

seu <- SetDimReduction(object = seu, reduction.type = "zinbwave", 
                       slot = "cell.embeddings",
                       new.data = reducedDim(zinbObj, "zinbwave"))

seu <- SetDimReduction(object = seu, reduction.type = "zinbwave", slot = "key",
                       new.data = "zinbwave")

resolution = 1
seu <- FindClusters(object = seu, reduction.type = "zinbwave", 
                    dims.use = 1:10, #this should match K
                    resolution = resolution, print.output = 0, save.SNN = TRUE)

so <- seu

z_cl <- so@ident
names(z_cl) <- paste0(names(so@ident))

table(z_cl)

##########_---------#############
z_cl <- as.factor(as.numeric(z_cl))
names(z_cl) <- colnames(sconeNorm)

write.table(data.frame(colnames(sconeNorm), paste0("zinbCl", z_cl)), sep="\t", quote=FALSE,
            file = paste0(outDir, "clustering/", exptStr, "_", resolution, "_zinb_SNN_RCOB_clusters.txt"))
# pairs(W[,1:5], cex = 0.3, col=colRKC[z_cl])
plot(tsne_zinb$Y, pch=19, cex = 0.4, col=colRKC[z_cl], xlab="TSNE 1", ylab="TSNE 2", main=paste0("t-SNE (ZINB-WaVE) 1KvarGenes, clustered by zinbwave\n resolution = ", resolution, ", varGenes = ", length(vargenes)))
legend("topleft", legend=levels(z_cl), fill=colRKC, cex=0.6)
plot(tsne_zinb$Y, pch=19, cex = 0.4, col=colb[batch], xlab="TSNE 1", ylab="TSNE 2", main="t-SNE (ZINB-WaVE) 1KvarGenes, clustered by zinbwave")
legend("bottomleft", legend=levels(batch), fill=colb, cex=0.6)

pdf(file=paste0(outDir, "viz/", exptStr, "_zinb_Seurat_tSNE.pdf"), title = "Zinb Seurat Clusters tSNE")
plot(tsne_zinb$Y, pch=19, cex = 0.4, col=colRKC[z_cl], xlab="TSNE 1", ylab="TSNE 2", main=paste0("t-SNE (ZINB-WaVE) 1KvarGenes, clustered by Seurat\n resolution = ", resolution, ", varGenes = ", length(vargenes)))
legend("topleft", legend=levels(z_cl), fill=colRKC, cex=0.5)
dev.off()
pdf(file=paste0(outDir, "viz/", exptStr, "_zinb_batch_tSNE.pdf"), title = "Zinb by batch tSNE")
plot(tsne_zinb$Y, pch=19, cex = 0.4, col=colb[batch], xlab="TSNE 1", ylab="TSNE 2", main="t-SNE (ZINB-WaVE) 1KvarGenes by batch")
legend("bottomleft", legend=levels(batch), fill=colb, cex=0.6)
dev.off()

zinbcl1 <- ClusterExperiment(sconeNorm, clusters=z_cl, transformation = log1p)
zinbcl1 <- makeDendrogram(zinbcl1, reduceMethod = "var", nDims=1000)
plotDendrogram(zinbcl1)

# zinbcl1 <- mergeClusters(zinbcl1, mergeMethod = "locfdr", cutoff = 0.05, plotInfo="mergeMethod")
colData(zinbcl1)$batch <- colData(filtered)$batch
#zm_cl <- zinbcl1@clusterMatrix[,1]

plotClusters(zinbcl1, sampleData = "batch", whichClusters = "all", colPalette = colRKC)

# plot(data.use, pch=19, cex=0.5, col=colRKC[z_cl])
pairs(W[,1:5], cex = 0.3, col=colRKC[z_cl])

pdf(file=paste0(outDir, "viz/", exptStr, "_zinb_Seurat_heatmap.pdf"), title = "Zinb Seurat Clusters heatmap")
clusterExperiment::plotHeatmap(log2(assay(zinbcl1)[obClusterGenes,]+1), clusterSamplesData = zinbcl1@dendro_samples, clusterFeatures=T, sampleData=data.frame(clusters=z_cl, batch=batch), clusterLegend=list(clusters=colRKC, batch=colb), main=paste("SNN/Seurat Zinb, select markers"))
dev.off()

top5pairs_zinb <- getBestFeatures(sconeNorm, cluster=z_cl, isCount=TRUE, contrastType = "Pairs", number=5)
pdf(file=paste0(outDir, "viz/", exptStr, "_heatmap_pairWise5_SNNzinb.pdf"), height=8.5, width=11)
plotHeatmap(log2(sconeNorm[unique(top5pairs_zinb$IndexInOriginal),]+1), clusterSamplesData = zinbcl1@dendro_samples, clusterFeatures=T, sampleData=data.frame(clusters=z_cl, batch=batch), clusterLegend=list(clusters=colRKC, batch=colb), main=paste("pairWise5_SNNzinb"))
dev.off()

```


###Now to do SNN on the first several PCs of the scone norm matrix
```{r SNNsconeNormPCA}
library(rARPACK)
# pcaScone <- fastpca(sconeNorm)
pca <- prcomp(t(sconeNorm), scale. = TRUE)
screeplot(pca, type = "lines", npcs = 50, main = "Expression PCA screeplot")

sceScone <- SingleCellExperiment(assays = list(counts = sconeNorm))
colData(sceScone) <- colData(filtered)
rowData(sceScone) <- rowData(filtered)
tsne_data <- Rtsne(pca$x[,1:30], pca = FALSE)
reducedDims(sceScone) <- SimpleList(PCA=pca$x, TSNE=tsne_data$Y)

seu2 <- CreateSeuratObject(raw.data = sconeNorm, min.cells = 1, min.genes = 1, 
     project = "RCOB")

seu2 <- SetDimReduction(object = seu2, reduction.type = "PCA", 
                       slot = "cell.embeddings",
                       new.data = reducedDim(sceScone, "PCA"))

seu2 <- SetDimReduction(object = seu2, reduction.type = "PCA", slot = "key",
                       new.data = "PCA")

resolution = 1
seu2 <- FindClusters(object = seu2, reduction.type = "PCA", 
                    dims.use = 1:30, #this should match dimensions
                    resolution = resolution, print.output = 0, save.SNN = TRUE)


s_cl <- seu2@ident
names(s_cl) <- paste0(names(seu2@ident))

table(s_cl)

s_cl <- as.factor(as.numeric(s_cl))
names(s_cl) <- colnames(sconeNorm)

write.table(data.frame(colnames(sconeNorm), paste0("sconeCl", s_cl)), sep="\t", quote=FALSE,
            file = paste0(outDir, "clustering/", exptStr, "_", resolution, "_zinb_SNN_clusters.txt"))
# pairs(W[,1:5], cex = 0.3, col=colRKC[s_cl])
plot(tsne_data$Y, pch=19, cex = 0.4, col=colRKC[s_cl], xlab="TSNE 1", ylab="TSNE 2", main=paste0(exptStr, " t-SNE (scone20PCA), clustered by Seurat\n resolution = ", resolution, ", varGenes = ", length(vargenes)))
legend("topleft", legend=levels(s_cl), fill=colRKC, cex=0.6)
plot(tsne_data$Y, pch=19, cex = 0.4, col=colb[batch], xlab="TSNE 1", ylab="TSNE 2", main=paste0(exptStr, "_t-SNE (scone20PCA) by batch"))
legend("topleft", legend=levels(batch), fill=colb, cex=0.6)

pdf(file=paste0(outDir, "viz/", exptStr, "_scone_Seurat_tSNE.pdf"), title = "Zinb Seurat Clusters tSNE")
plot(tsne_data$Y, pch=19, cex = 0.4, col=colRKC[s_cl], xlab="TSNE 1", ylab="TSNE 2", main=paste0(exptStr, " t-SNE (scone20PCA), clustered by Seurat\n resolution = ", resolution, ", varGenes = ", length(vargenes)))
legend("topleft", legend=levels(s_cl), fill=colRKC, cex=0.6)
dev.off()
pdf(file=paste0(outDir, "viz/", exptStr, "_scone_batch_tSNE.pdf"), title = "Zinb by batch tSNE")
plot(tsne_data$Y, pch=19, cex = 0.4, col=colb[batch], xlab="TSNE 1", ylab="TSNE 2", main=paste0(exptStr, "_t-SNE (scone20PCA) by batch"))
legend("topleft", legend=levels(batch), fill=colb, cex=0.6)
dev.off()

sconeCE <- ClusterExperiment(sconeNorm, clusters=s_cl, transformation = log1p)
sconeCE <- makeDendrogram(sconeCE, reduceMethod = "var", nDims=1000)
plotDendrogram(sconeCE)

# sconeCE <- mergeClusters(sconeCE, mergeMethod = "locfdr", cutoff = 0.05, plotInfo="mergeMethod")
colData(sconeCE)$batch <- batch
#zm_cl <- sconeCE@clusterMatrix[,1]

plotClusters(sconeCE, sampleData = "batch", whichClusters = "all", colPalette = colRKC)

# plot(data.use, pch=19, cex=0.5, col=colRKC[z_cl])
pairs(pca_sconeNorm[,1:5], cex = 0.3, col=colRKC[s_cl])

pdf(file=paste0(outDir, "viz/", exptStr, "_scone_Seurat_heatmap.pdf"), title = "scone Seurat Clusters heatmap")
clusterExperiment::plotHeatmap(log2(assay(sconeCE)[obClusterGenes,]+1), clusterSamplesData = sconeCE@dendro_samples, clusterFeatures=T, sampleData=data.frame(clusters=s_cl, batch=batch), clusterLegend=list(clusters=colRKC, batch=colb), main=paste("SNN/Seurat scone, select markers"))
dev.off()

top5pairs_scone <- getBestFeatures(sconeNorm, cluster=s_cl, isCount=TRUE, contrastType = "Pairs", number=5)
pdf(file=paste0(outDir, "viz/", exptStr, "_heatmap_pairWise5_SNNscone.pdf"), height=8.5, width=11)
plotHeatmap(log2(sconeNorm[unique(top5pairs_scone$IndexInOriginal),]+1), clusterSamplesData = sconeCE@dendro_samples, clusterFeatures=T, sampleData=data.frame(clusters=z_cl, batch=batch), clusterLegend=list(clusters=colRKC, batch=colb), main=paste("pairWise5_SNNscone"))
dev.off()
```

```{r sconeSeuratDE}

oneVallDE1K <- getBestFeatures(sconeNorm, cluster=s_cl, isCount=TRUE, contrastType = "OneAgainstAll", number=1000)
write.table(oneVallDE1K, sep="\t", quote=FALSE,
            file = paste0(outDir, "DE/", exptStr, "_sconeSNN_oneVallDE1K.txt"))

pairwiseDE500 <- getBestFeatures(sconeNorm, cluster=s_cl, isCount=TRUE, contrastType = "Pairs", number=500)
write.table(pairwiseDE500, sep="\t", quote=FALSE,
            file = paste0(outDir, "DE/", exptStr, "_sconeSNN_pairwiseDE500.txt"))



```



```{r SNNsconeTSNE}

#####-----######------######

seu3 <- CreateSeuratObject(raw.data = sconeNorm, min.cells = 1, min.genes = 1, 
     project = "RCOB")

seu3 <- SetDimReduction(object = seu3, reduction.type = "TSNE", 
                       slot = "cell.embeddings",
                       new.data = reducedDim(sceScone, "TSNE"))

seu3 <- SetDimReduction(object = seu3, reduction.type = "TSNE", slot = "key",
                       new.data = "TSNE")

resolution = 0.6
seu3 <- FindClusters(object = seu3, reduction.type = "TSNE", 
                    dims.use = 1:2, #this should match dimensions
                    resolution = resolution, print.output = 0, save.SNN = TRUE)


s_cl2 <- seu3@ident
names(s_cl2) <- paste0(names(seu3@ident))

table(s_cl2)

s_cl2 <- as.factor(as.numeric(s_cl2))
names(s_cl2) <- colnames(sconeNorm)

write.table(data.frame(colnames(sconeNorm), paste0("sconeTSNE", s_cl2)), sep="\t", quote=FALSE,
            file = paste0(outDir, "clustering/", exptStr, "_", resolution, "_sconeTSNE_SNN_clusters.txt"))
# pairs(W[,1:5], cex = 0.3, col=colRKC[s_cl])
plot(tsne_data$Y, pch=19, cex = 0.4, col=colRKC[s_cl2], xlab="TSNE 1", ylab="TSNE 2", main=paste0(exptStr, " t-SNE, clustered by Seurat on t-SNE\n resolution = ", resolution, ", varGenes = ", length(vargenes)))
legend("bottomleft", legend=levels(s_cl2), fill=colRKC, cex=0.55)
plot(tsne_data$Y, pch=19, cex = 0.4, col=colb[batch], xlab="TSNE 1", ylab="TSNE 2", main=paste0(exptStr, "_t-SNE (scone20PCA) by batch"))
legend("topleft", legend=levels(batch), fill=colb, cex=0.6)

pdf(file=paste0(outDir, "viz/", exptStr, "_sconetSNE_SeuratTSNE_tSNE.pdf"), title = "sconeTSNE Seurat Clusters tSNE")
plot(tsne_data$Y, pch=19, cex = 0.4, col=colRKC[s_cl2], xlab="TSNE 1", ylab="TSNE 2", main=paste0(exptStr, " t-SNE, clustered by Seurat on t-SNE\n resolution = ", resolution, ", varGenes = ", length(vargenes)))
legend("topleft", legend=levels(s_cl2), fill=colRKC, cex=0.6)
dev.off()
pdf(file=paste0(outDir, "viz/", exptStr, "_scone20PCA_batch_tSNE.pdf"), title = "scone by batch tSNE")
plot(tsne_data$Y, pch=19, cex = 0.4, col=colb[batch], xlab="TSNE 1", ylab="TSNE 2", main=paste0(exptStr, "_t-SNE (scone20PCA) by batch"))
legend("topleft", legend=levels(batch), fill=colb, cex=0.6)
dev.off()


sconeCE <- ClusterExperiment(sconeNorm, clusters=s_cl2, transformation = log1p)
sconeCE <- makeDendrogram(sconeCE, reduceMethod = "var", nDims=1000)
plotDendrogram(sconeCE)
pdf(file=paste0(outDir, "viz/", exptStr, "_scone_SeuratTSNE_heatmap.pdf"), title = "scone Seurat Clusters heatmap")
clusterExperiment::plotHeatmap(log2(assay(sconeCE)[obClusterGenes,]+1), clusterSamplesData = sconeCE@dendro_samples, clusterFeatures=T, sampleData=data.frame(clusters=s_cl2, batch=batch), clusterLegend=list(clusters=colRKC, batch=colb), main=paste("SNN/Seurat sconeTSNE, select markers"))
dev.off()
```

```{r sconeTSNESeuratDE}

oneVallDE1K <- getBestFeatures(sconeNorm, cluster=s_cl2, isCount=TRUE, contrastType = "OneAgainstAll", number=1000)
write.table(oneVallDE1K, sep="\t", quote=FALSE,
            file = paste0(outDir, "DE/", exptStr, "_sconeTSNE_SNN_oneVallDE1K.txt"))

pairwiseDE500 <- getBestFeatures(sconeNorm, cluster=s_cl2, isCount=TRUE, contrastType = "Pairs", number=500)
write.table(pairwiseDE500, sep="\t", quote=FALSE,
            file = paste0(outDir, "DE/", exptStr, "_sconeTSNE_SNN_pairwiseDE500.txt"))

```

ce <- ClusterExperiment(sconeNorm, clusters=s_cl2, transformation = log1p)
ce <- makeDendrogram(ce, reduceMethod = "var", nDims=1000)
plotDendrogram(ce)

# sconecl1 <- mergeClusters(sconecl1, mergeMethod = "locfdr", cutoff = 0.05, plotInfo="mergeMethod")
colData(ce)$batch <- batch
#zm_cl <- sconecl1@clusterMatrix[,1]

plotClusters(ce, sampleData = "batch", whichClusters = "all", colPalette = colRKC)

# plot(data.use, pch=19, cex=0.5, col=colRKC[s_cl2])
pairs(W[,1:5], cex = 0.3, col=colRKC[s_cl2])

pdf(file=paste0(outDir, "viz/", exptStr, "_sconeTSNE_Seurat_heatmap.pdf"), title = "sconeTSNE Seurat Clusters heatmap")
clusterExperiment::plotHeatmap(log2(assay(ce)[obClusterGenes,]+1), clusterSamplesData = ce@dendro_samples, clusterFeatures=T, sampleData=data.frame(clusters=s_cl2, batch=batch), clusterLegend=list(clusters=colRKC, batch=colb), main=paste("SNN/Seurat sconeTSNE, select markers"))
dev.off()

top5pairs_sconeTSNE <- getBestFeatures(sconeNorm, cluster=s_cl2, isCount=TRUE, contrastType = "Pairs", number=5)
pdf(file=paste0(outDir, "viz/", exptStr, "_heatmap_pairWise5_SNNsconeTSNE.pdf"), height=8.5, width=11)
clusterExperiment::plotHeatmap(log2(sconeNorm[unique(top5pairs_sconeTSNE$IndexInOriginal),]+1), clusterSamplesData = ce@dendro_samples, clusterFeatures=T, sampleData=data.frame(clusters=s_cl2, batch=batch), clusterLegend=list(clusters=colRKC, batch=colb), main=paste("pairWise5_SNNsconeTSNE"))
dev.off()

```







```{r vizSeuratScone}
top10_scone50 <- getBestFeatures(sconeNorm, cluster=scone50_cl, isCount=TRUE, contrastType = "OneAgainstAll", number=10)
top500_scone50 <- getBestFeatures(sconeNorm, cluster=scone50_cl, isCount=TRUE, contrastType = "OneAgainstAll", number=500)
write.table(top500_scone50, sep="\t", quote=FALSE,
            file = paste0(outDir, "DE/", exptStr, "_scone50SNN_sconeNorm_oneVallDE500.txt"))

top5pairs_scone50 <- getBestFeatures(sconeNorm, cluster=scone50_cl, isCount=TRUE, contrastType = "Pairs", number=5)
top100pairs_scone50 <- getBestFeatures(sconeNorm, cluster=scone50_cl, isCount=TRUE, contrastType = "Pairs", number=100)
write.table(top100pairs_scone50, sep="\t", quote=FALSE,
            file = paste0(outDir, "DE/", exptStr, "_scone50_clSNN_sconeNorm_pairsDE100.txt"))

pdf(file=paste0(outDir, "viz/", exptStr, "_heatmap_oneVall10_SNNscone50_sconeNorm.pdf"), height=8.5, width=11)
clusterExperiment::plotHeatmap(log2(sconeNorm[unique(top10_scone$IndexInOriginal),]+1), clusterSamplesData = sconeCl1@dendro_samples, clusterFeatures=T, sampleData=data.frame(clusters=z_cl, batch=colData(filtered)[,1]), clusterLegend=list(clusters=colRKC, batch=colb), main=paste(exptStr, "_oneVall10_SNNscone50_sconeNorm")) 
dev.off()

pdf(file=paste0(outDir, "viz/", exptStr, "_heatmap_pairWise5_SNNscone50_sconeNorm.pdf"), height=8.5, width=11)
plotHeatmap(log2(sconeNorm[unique(top5pairs_scone50$IndexInOriginal),]+1), clusterSamplesData = scone50cl1@dendro_samples, clusterFeatures=T, sampleData=data.frame(clusters=z_cl, batch=colData(filtered)[,1]), clusterLegend=list(clusters=colRKC, batch=colb), main=paste("pairWise5_SNNscone501Kvar_sconeNorm"))
dev.off()


```


#Everything below this point is older code.






```{r seurat_zinb2, echo=FALSE, eval=FALSE}

library(dplyr)
library(Matrix)

so <- CreateSeuratObject(raw.data = sconeNorm, min.cells = 1, min.genes = 1, 
    project = "regen1")


## Build SNN
k.param = 10
k.scale = 10
n.cells = 942
data.use <- getW(zinb)
my.knn <- FNN::get.knn(as.matrix(data.use), k = min(k.scale * k.param, n.cells - 1))
nn.ranked <- cbind(1:n.cells, my.knn$nn.index[, 1:(k.param-1)])
nn.large <- my.knn$nn.index

w <- Seurat:::CalcSNNSparse(cell.names = colnames(filtCounts),
                            k.param = k.param,
                            nn.large = nn.large,
                            nn.ranked = nn.ranked,
                            prune.SNN = 1/15,
                            print.output = FALSE)

so@snn <- w

## Run modularity clustering
#resList <- c(1.0,1.1,1.2,1.3,1.4,1.5,1.6,1.7,1.8,1.9,2.0,2.1,2.2,2.3,2.4,2.5,2.6,2.7,2.8,2.9,3.0)
resList <- c(0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1.0,1.1,1.2,1.3,1.4,1.5,1.6,1.7,1.8,1.9,2.0)
#cors <- lapply(1:5, function(i) abs(cor(pca_sconeNorm$x[,i], qc, method="spearman")))
resolution <- 1
so <- Seurat:::RunModularityClustering(object = so, SNN = w, 
                                         modularity = 1, resolution = resolution,
                                         algorithm = 1, n.start = 100, 
                                         n.iter = 10, random.seed = 0,
                                         print.output = FALSE, temp.file.location = NULL)
so <- Seurat:::GroupSingletons(so, so@snn)
name <- paste("res.", resolution, sep = "")
so <- StashIdent(so, name)

z_cl0 <- so@ident
names(z_cl0) <- paste0(names(so@ident))
z_cl0 <- as.matrix(z_cl0)
SNNcl <- lapply(resList, function(i) {
  resolution <- i
  so <- Seurat:::RunModularityClustering(object = so, SNN = w, 
                                         modularity = 1, resolution = resolution,
                                         algorithm = 1, n.start = 100, 
                                         n.iter = 10, random.seed = 0,
                                         print.output = FALSE, temp.file.location = NULL)
  so <- Seurat:::GroupSingletons(so, so@snn)
  name <- paste("res.", resolution, sep = "")
  so <- StashIdent(so, name)
  z_cl1 <- so@ident
  names(z_cl1) <- paste0(names(so@ident))
  #return(z_cl)
  z_cl0 <<- cbind(z_cl0, z_cl1)
})

z_clCombo <- z_cl0[,-1]
colnames(z_clCombo) <- resList

zinbce <- clusterExperiment(sconeNorm, clusters=z_clCombo, clusterTypes = as.character(resList),
                       transformation = log1p)
#colnames(zinbce@clusterMatrix) <- as.character(resList)
plotClusters(zinbce)
zinbcl <- clusterMatrix(zinbce)
zinbcm <- combineMany(zinbcl, proportion = .5, minSize = 10)
zcm_cl <- as.factor(zinbcm$clustering)
plotClusters(cbind(zinbcm$clustering, zinbcl))
zinbnew <- clusterExperiment(sconeNorm, clusters=zinbcm$clustering, transformation = log1p)
zinbnew <- makeDendrogram(zinbnew, dimReduce="var", ndims=1000)
plotDendrogram(zinbnew)

zinbnew <- mergeClusters(zinbnew, mergeMethod = "locfdr", cutoff = 0.08, plotInfo="mergeMethod")
rm(zinbce)
colData(zinbnew)$batch <- colData(filtered)$batch

plotClusters(zinbnew, sampleData = "batch", whichClusters = "all", colPalette = colRKC)
zcmm_cl <- zinbnew@clusterMatrix[,1]
zcmm_cl <- factor(clusterMatrixNamed(zinbnew)[,"mergeClusters"],
                      levels=clusterLegend(zinbnew)[["mergeClusters"]][, "name"])
names(zcmm_cl) <- colnames(zinbnew)
wh_rm <- which(zcmm_cl=="-1")
# pca2 <- prcomp(t(lognorm[,-twh_rm]))

plot(tsne_zinb$Y, pch=19, cex = 0.4, col=colRKC[z_cl[-wh_rm]], xlab="TSNE 1", ylab="TSNE 2", main="t-SNE (ZINB-WaVE) 1KvarGenes, clustered by zinbwave\n combineMany, resolution range")
legend("bottomleft", legend=levels(z_cl[-wh_rm]), fill=colRKC, cex=0.6)

plot(tsne_zinb$Y, pch=19, cex = 0.4, col=cole[batch], xlab="TSNE 1", ylab="TSNE 2", main="t-SNE (ZINB-WaVE) 1KvarGenes, clustered by zinbwave")
legend("bottomleft", legend=levels(batch), fill=cole, cex=0.6)

plot(data.use, pch=19, cex=0.5, col=colRKC[zcmm_cl])
pairs(W[,1:5], cex = 0.3, col=colRKC[zcmm_cl])
```

```{r zinbClusterExperimentObject, eval=FALSE}

zinbce <- clusterExperiment(sconeNorm, clusters=cbind(as.character(zm_cl),      as.character(z_cl),as.character(zcmm_cl), as.character(zcm_cl)), clusterTypes = c("Merged", "res2", "CMmerged", "CM"),
                       transformation = log1p)
colnames(ce@clusterMatrix) <- c("zinbMerged", "zinbRes2", "zinbCMmerged", "zinbCM")

```

###SNN clustering on the scone norm, first 50 PCs
```{r seurat_sconenorm_pca50, echo=FALSE, eval=TRUE}

vargenes <- names(vars)[1:1000]
pca_sconeNorm <- fastpca(log2(sconeNorm[vargenes,] + 0.1))
#pca_sconeNorm <- fastpca(log2(sconeNorm + 0.1))
so <- CreateSeuratObject(raw.data = sconeNorm, min.cells = 1, min.genes = 1, 
    project = "RCOB")

## Build SNN
k.param = 10
k.scale = 10
n.cells = 942
data.use <- pca_sconeNorm[,1:50]
my.knn <- FNN::get.knn(as.matrix(data.use), k = min(k.scale * k.param, n.cells - 1))
nn.ranked <- cbind(1:n.cells, my.knn$nn.index[, 1:(k.param-1)])
nn.large <- my.knn$nn.index

w <- Seurat:::CalcSNNSparse(cell.names = colnames(filtCounts),
                            k.param = k.param,
                            nn.large = nn.large,
                            nn.ranked = nn.ranked,
                            prune.SNN = 1/15,
                            print.output = FALSE)

so@snn <- w

## Run modularity clustering
resolution <- 0.6
so <- Seurat:::RunModularityClustering(object = so, SNN = w, 
                                         modularity = 1, resolution = resolution,
                                         algorithm = 1, n.start = 100, 
                                         n.iter = 10, random.seed = 0,
                                         print.output = FALSE, temp.file.location = NULL)
so <- Seurat:::GroupSingletons(so, so@snn)
name <- paste("res.", resolution, sep = "")
so <- StashIdent(so, name)

scone50_cl <- so@ident
names(scone50_cl) <- paste0(names(so@ident), "-1")

table(scone50_cl)
table(scone50_cl, z_cl)

#plot(data.use, pch=19, col=pal[scone50_cl])
colnames(data.use) <- paste0("PCA", 1:NCOL(pca_sconeNorm))
rownames(data.use) <- names(scone50_cl)
write.table(data.frame(data.use, paste0("sconeNorm_PCA50cl", scone50_cl)), sep="\t", quote=FALSE,
            file = "sconeNorm_pca50_SNN_RCOB_clusters.txt")
#pairs(pca_sconeNorm[,1:5], cex = 0.3, col=pal[scone50_cl])

tsne_scone <- Rtsne(pca_sconeNorm[,1:20], pca=FALSE, max_iter = 5000)
plot(tsne_scone$Y, pch=19, cex = 0.4, col=colRKC[scone50_cl], xlab="TSNE 1", ylab="TSNE 2", main=paste0("t-SNE (sconeNorm PCA20), clustered by sconeNorm PCA50\n resolution = ", resolution, ", varGenes = ", length(vargenes)))
legend("bottomleft", legend=levels(scone50_cl), fill=colRKC, cex=0.5)
plot(tsne_scone$Y, pch=19, cex = 0.4, col=colb[batch], xlab="TSNE 1", ylab="TSNE 2", main="t-SNE (sconeNorm PCA20), clustered by sconeNorm PCA50")
legend("bottomleft", legend=levels(batch), fill=colb, cex=0.5)
```

#### Visually comparing the different clusterings from Seurat (zinb (z_cl), sconeNorm 50 PCs (scone50_cl)
```{r clusterComparison, echo=FALSE}


table(z_cl, scone50_cl)


plot(tsne_zinb$Y, pch=19, col=colRKC[z_cl], cex=0.4, xlab="TSNE 1", ylab="TSNE 2", main="t-SNE (ZINB-WaVE_1Kvar), clustered by zinbwave1Kvar")

plot(tsne_scone$Y, pch=19, col=colRKC[z_cl], cex=0.4, xlab="TSNE 1", ylab="TSNE 2", main="t-SNE (sconeNormPCA, 50PCs), clustered by zinbwave1Kvar")


plot(tsne_scone$Y, pch=19, col=colRKC[scone50_cl], cex=0.4, xlab="TSNE 1", ylab="TSNE 2", main="t-SNE (sconeNormPCA, 50PCs), clustered by sconeNormPCA 50PCs")

plot(tsne_zinb$Y, pch=19, col=colRKC[scone50_cl], cex=0.4, xlab="TSNE 1", ylab="TSNE 2", main="t-SNE (ZINB-WaVE_1Kvar), clustered by sconeNormPCA 50PCs")

```

```{r savingData, echo=FALSE}
#save(filtCounts, tsne_data, tsne_data10, z_cl, z1k_cl, scone50_cl, sconen_pca10_cl, sumn_pca_cl, sumn_pca10_cl, tsne_zinb, data.pca, data.zinb, zinb, filtered, pca_sumNorm, pca_sumNorm1000, sumNorm, sumNorm1000, pca_sconeNorm, sconeNorm, vargenes, batch, file="regen1_toStartClusterInterpretation.Rda")
save(pca, qcpca, file="dataForQCanalysis.Rda")

batch <- colData(filtered)$batch
expt <- colData(filtered)$expt

# save(qcpca, poscon, negconeval, negconruv, filtCounts, z_cl, scone50_cl, filtered, pca_sconeNorm, sconeNorm, logSconeNorm, vargenes, zinb, tsne_scone, tsne_zinb, expt, batch,  tsne_data, W, cole, colb, logfiltCounts, file="RCOB_forClusterInterpretation.Rda")
# rm(zinbcl1, so, se, scone_obj, scone_obj2, scone_obj3, scone_obj4, runQCmetrics, run_zinbwave, run_scone, ribo_idx, mito_idx, ribo_pct, mito_pct, mfilt, is_common, is_quality, cors, d, fig_barplot, fig_data, fig_pca, fig_tsne)
```


#####Make clusterExperiment object and use to compare clusterings, DE, etc.
```{r preparationClusterExperiment}

z_cl <- as.factor(as.numeric(z_cl))
names(z_cl) <- colnames(sconeNorm)

scone50_cl <- as.factor(as.numeric(scone50_cl))
names(scone50_cl) <- colnames(sconeNorm)

ce <- clusterExperiment(filtered, clusters=cbind(as.character(z_cl),
                                          as.character(scone50_cl)), clusterTypes = c("zinb1Kvar", "scone50PC"),
                       transformation = log1p)
colnames(ce@clusterMatrix) <- c("zinb1Kvar", "scone50PC")
assay(ce) <- sconeNorm

ce <- makeDendrogram(ce, dimReduce="PCA", ndims=10)

plotClusters(ce)
plotDendrogram(ce)




# ce <- clusterExperiment(filtered, clusters=cbind(as.character(z_cl),
#                                           as.character(scone50_cl)), clusterTypes = c("zinb1Kvar", "scone50PC"),
#                        transformation = log1p)
# colnames(ce@clusterMatrix) <- c("zinb1Kvar", "scone50PC")
# assay(ce) <- sconeNorm
# 
# ce <- makeDendrogram(ce, dimReduce="PCA", ndims=10)
```

#####Visualize gene expression for the candidate normalization/clusterings
```{r visualizeGeneExpressionByCluster}

###visualizing gene expression by cluster, using SCONE norm counts

############-----------First, with SNN of Zinb1Kvar
library(ggplot2)
theme_set(theme_classic())

markers <- c("Thy1", "Gad1", "Fxyd6", "Fxyd7", "Bcl2", "Pcp4", "Meis2", "Kcna2", "Gabra1", "Grin2b", "Dnajc2", "Vps8", "Calb2", "Fstl5", "Camk2n1", "Gls", "Reln", "Slc17a7", "Fst", "Coch", "tdTomato", "CreER")
fig_data <- data.frame(t(log2(sconeNorm[markers,]+1)),
                         TSNE1 = tsne_zinb$Y[,1], TSNE2 = tsne_zinb$Y[,2],
                         Cluster = as.factor(z_cl), qc)

ggplot(fig_data, aes(x = TSNE1, y = TSNE2, color = log10_total_counts)) +
  geom_point() + scale_color_continuous(low = "blue", high = "yellow")

ggplot(fig_data, aes(x = TSNE1, y = TSNE2, color = ribo_pct)) +
  geom_point() + scale_color_continuous(low = "blue", high = "yellow")

ggplot(aes(TSNE1, TSNE2, colour=Cluster), data = fig_data) + 
 geom_point() + scale_colour_manual(values = colRKC)

ggplot(aes(TSNE1, TSNE2, colour=Gad1), data = fig_data) + 
 geom_point() + scale_colour_gradient(low="blue", high="yellow")

ggplot(aes(TSNE1, TSNE2, colour=Camk2n1), data = fig_data) + 
 geom_point() + scale_colour_gradient(low="blue", high="yellow")

ggplot(aes(TSNE1, TSNE2, colour=Meis2), data = fig_data) + 
 geom_point() + scale_colour_gradient(low="blue", high="yellow")

ggplot(aes(TSNE1, TSNE2, colour=Pcp4), data = fig_data) + 
 geom_point() + scale_colour_gradient(low="blue", high="yellow")

ggplot(aes(TSNE1, TSNE2, colour=Thy1), data = fig_data) + 
 geom_point() + scale_colour_gradient(low="blue", high="yellow")

ggplot(aes(TSNE1, TSNE2, colour=Reln), data = fig_data) + 
 geom_point() + scale_colour_gradient(low="blue", high="yellow")

ggplot(aes(TSNE1, TSNE2, colour=Fxyd6), data = fig_data) + 
 geom_point() + scale_colour_gradient(low="blue", high="yellow")

ggplot(aes(TSNE1, TSNE2, colour=Fxyd7), data = fig_data) + 
 geom_point() + scale_colour_gradient(low="blue", high="yellow")

ggplot(aes(TSNE1, TSNE2, colour=Bcl2), data = fig_data) + 
 geom_point() + scale_colour_gradient(low="blue", high="yellow")

ggplot(aes(TSNE1, TSNE2, colour=Kcna2), data = fig_data) + 
 geom_point() + scale_colour_gradient(low="blue", high="yellow")

ggplot(aes(TSNE1, TSNE2, colour=Fstl5), data = fig_data) + 
 geom_point() + scale_colour_gradient(low="blue", high="yellow")

ggplot(aes(TSNE1, TSNE2, colour=Fst), data = fig_data) + 
 geom_point() + scale_colour_gradient(low="blue", high="yellow")

ggplot(aes(TSNE1, TSNE2, colour=Gls), data = fig_data) + 
 geom_point() + scale_colour_gradient(low="blue", high="yellow")

ggplot(aes(TSNE1, TSNE2, colour=Slc17a7), data = fig_data) + 
 geom_point() + scale_colour_gradient(low="blue", high="yellow")

ggplot(aes(TSNE1, TSNE2, colour=Coch), data = fig_data) + 
 geom_point() + scale_colour_gradient(low="blue", high="yellow")

ggplot(aes(TSNE1, TSNE2, colour=tdTomato), data = fig_data) + 
 geom_point() + scale_colour_gradient(low="blue", high="yellow")

ggplot(aes(TSNE1, TSNE2, colour=CreER), data = fig_data) + 
 geom_point() + scale_colour_gradient(low="blue", high="yellow")


########################################
########################################

#####saving relevant files for Rebecca so she can plot tSNEs and heatmaps of relevant genes
save(sconeNorm, ce, vargenes, batch, expt, resolution, tsne_scone, scone50_cl, z_cl, qc, file="forGeneExpression_tSNEs.Rda")

############-----------with SNN of 50 PCs sconeNorm

theme_set(theme_classic())
fig_data <- data.frame(t(log2(sconeNorm[markers,]+1)),
                         TSNE1 = tsne_scone$Y[,1], TSNE2 = tsne_scone$Y[,2],
                         Cluster = scone50_cl, qc)

ggplot(aes(TSNE1, TSNE2, colour=Cluster), data = fig_data) + 
 geom_point() + scale_colour_manual(values = colRKC)

ggplot(fig_data, aes(x = TSNE1, y = TSNE2, color = log10_total_counts)) +
  geom_point() + scale_color_continuous(low = "blue", high = "yellow")

ggplot(fig_data, aes(x = TSNE1, y = TSNE2, color = ribo_pct)) +
  geom_point() + scale_color_continuous(low = "blue", high = "yellow")

ggplot(aes(TSNE1, TSNE2, colour=Gad1), data = fig_data) + 
 geom_point() + scale_colour_gradient(low="blue", high="yellow")

ggplot(aes(TSNE1, TSNE2, colour=Camk2n1), data = fig_data) + 
 geom_point() + scale_colour_gradient(low="blue", high="yellow")

ggplot(aes(TSNE1, TSNE2, colour=Meis2), data = fig_data) + 
 geom_point() + scale_colour_gradient(low="blue", high="yellow")

ggplot(aes(TSNE1, TSNE2, colour=Pcp4), data = fig_data) + 
 geom_point() + scale_colour_gradient(low="blue", high="yellow")

ggplot(aes(TSNE1, TSNE2, colour=Thy1), data = fig_data) + 
 geom_point() + scale_colour_gradient(low="blue", high="yellow")

ggplot(aes(TSNE1, TSNE2, colour=Reln), data = fig_data) + 
 geom_point() + scale_colour_gradient(low="blue", high="yellow")

ggplot(aes(TSNE1, TSNE2, colour=Fxyd6), data = fig_data) + 
 geom_point() + scale_colour_gradient(low="blue", high="yellow")

ggplot(aes(TSNE1, TSNE2, colour=Fxyd7), data = fig_data) + 
 geom_point() + scale_colour_gradient(low="blue", high="yellow")

ggplot(aes(TSNE1, TSNE2, colour=Bcl2), data = fig_data) + 
 geom_point() + scale_colour_gradient(low="blue", high="yellow")

ggplot(aes(TSNE1, TSNE2, colour=Kcna2), data = fig_data) + 
 geom_point() + scale_colour_gradient(low="blue", high="yellow")

ggplot(aes(TSNE1, TSNE2, colour=Fstl5), data = fig_data) + 
 geom_point() + scale_colour_gradient(low="blue", high="yellow")

ggplot(aes(TSNE1, TSNE2, colour=Fst), data = fig_data) + 
 geom_point() + scale_colour_gradient(low="blue", high="yellow")

ggplot(aes(TSNE1, TSNE2, colour=Gls), data = fig_data) + 
 geom_point() + scale_colour_gradient(low="blue", high="yellow")

ggplot(aes(TSNE1, TSNE2, colour=Slc17a7), data = fig_data) + 
 geom_point() + scale_colour_gradient(low="blue", high="yellow")

ggplot(aes(TSNE1, TSNE2, colour=Coch), data = fig_data) + 
 geom_point() + scale_colour_gradient(low="blue", high="yellow")




##################################################

```


####Visualization of gene expression by heatmaps for each clustering
```{r heatmapsDifferentClusterings}

plot(tsne_zinb$Y, pch=19, cex = 0.4, col=colRKC[z_cl], xlab="TSNE 1", ylab="TSNE 2", main=paste0("t-SNE (SNN_zinb1Kvar)\n resolution = ", resolution))
legend("topleft", levels(z_cl), fill=colRKC, cex=.5)
pdf(file="output/viz/tSNE_SNN_zinb1KvarCl.pdf")
plot(tsne_zinb$Y, pch=19, cex = 0.4, col=colRKC[z_cl], xlab="TSNE 1", ylab="TSNE 2", main=paste0("t-SNE (SNN_zinb1Kvar)\n resolution = ", resolution))
legend("topleft", levels(z_cl), fill=colRKC, cex=.5)
dev.off()

primaryClusterIndex(ce) <- 1
ce <- makeDendrogram(ce, dimReduce="PCA", ndims=10)
plotDendrogram(ce)

plotHeatmap(log2(assay(ce)[markers,]+1), clusterSamplesData = ce@dendro_samples, clusterFeatures=T, sampleData=data.frame(clusters=z_cl, clusters2=scone50_cl, batch=colData(ce)[,1]), clusterLegend=list(clusters=colRKC, clusters2=colRKC, batch=colb), main=paste("SNN Zinb1KvarCl, select markers"))
pdf(file="output/viz/heatmap_SNN_zinb1KvarCl_markers.pdf", height=8.5, width=11)
plotHeatmap(log2(assay(ce)[markers,]+1), clusterSamplesData = ce@dendro_samples, clusterFeatures=T, sampleData=data.frame(clusters=z_cl, clusters2=scone50_cl, batch=colData(ce)[,1]), clusterLegend=list(clusters=colRKC, clusters2=colRKC, batch=colb), main=paste("SNN Zinb1KvarCl, select markers"))
dev.off()

#################
plot(tsne_scone$Y, pch=19, cex = 0.4, col=colRKC[scone50_cl], xlab="TSNE 1", ylab="TSNE 2", main=paste0("t-SNE (SNN_scone50)\n resolution = ", resolution))
legend("topleft", levels(scone50_cl), fill=colRKC, cex=.5)
pdf(file="output/viz/tSNE_SNN_scone50PCCl.pdf")
plot(tsne_scone$Y, pch=19, cex = 0.4, col=colRKC[scone50_cl], xlab="TSNE 1", ylab="TSNE 2", main=paste0("t-SNE (SNN_scone50)\n resolution = ", resolution))
legend("topleft", levels(scone50_cl), fill=colRKC, cex=.5)
dev.off()

primaryClusterIndex(ce) <- 2
ce <- makeDendrogram(ce, dimReduce="PCA", ndims=10)
plotDendrogram(ce)

plotHeatmap(log2(assay(ce)[markers,]+1), clusterSamplesData = ce@dendro_samples, clusterFeatures=T, sampleData=data.frame(clusters=scone50_cl, clusters2=z_cl, batch=colData(ce)[,1]), clusterLegend=list(clusters=colRKC, clusters2=colRKC, batch=colb), main=paste("SNN scone50PCsCl, select markers"))
pdf(file="output/viz/heatmap_SNN_scone50PC_markers.pdf", height=8.5, width=11)
plotHeatmap(log2(assay(ce)[markers,]+1), clusterSamplesData = ce@dendro_samples, clusterFeatures=T, sampleData=data.frame(clusters=scone50_cl, clusters2=z_cl, batch=colData(ce)[,1]), clusterLegend=list(clusters=colRKC, clusters2=colRKC, batch=colb), main=paste("SNN scone50PCsCl, select markers"))
dev.off()

```

###DE genes for two selected clusterings using Limma Voom on sconeNormalized counts
```{r DEbyCluster}

#####-----Using sconeNorm counts for all DE
#####-----First, DE by cluster for SNN clustering of ZINBWaVE 1Kvar W

primaryClusterIndex(ce) <- 1
ce <- makeDendrogram(ce, dimReduce="PCA", ndims=10)
plotDendrogram(ce)

top5_zinb <- getBestFeatures(sconeNorm, cluster=z_cl, isCount=TRUE, contrastType = "OneAgainstAll", number=5)
top10_zinb <- getBestFeatures(sconeNorm, cluster=z_cl, isCount=TRUE, contrastType = "OneAgainstAll", number=10)
top1K_zinb <- getBestFeatures(sconeNorm, cluster=z_cl, isCount=TRUE, contrastType = "OneAgainstAll", number=1000)
write.table(top1K_zinb, sep="\t", quote=FALSE,
            file = paste0(outDir, "DE/", exptStr, "_zinbSNN_sconeNorm_oneVallDE1K.txt"))

top5pairs_zinb <- getBestFeatures(sconeNorm, cluster=z_cl, isCount=TRUE, contrastType = "Pairs", number=5)
top25pairs_zinb <- getBestFeatures(sconeNorm, cluster=z_cl, isCount=TRUE, contrastType = "Pairs", number=25)
write.table(top25pairs_zinb, sep="\t", quote=FALSE,
            file = paste0(outDir, "DE/", exptStr, "_zinbSNN_sconeNorm_pairsDE25.txt"))
top1Kpairs_zinb <- getBestFeatures(sconeNorm, cluster=z_cl, isCount=TRUE, contrastType = "Pairs", number=1000)
write.table(top1Kpairs_zinb, sep="\t", quote=FALSE,
            file = paste0(outDir, "DE/", exptStr, "_zinbSNN_sconeNorm_pairsDE1K.txt"))
top500pairs_zinb <- getBestFeatures(sconeNorm, cluster=z_cl, isCount=TRUE, contrastType = "Pairs", number=500)
write.table(top500pairs_zinb, sep="\t", quote=FALSE,
            file = paste0(outDir, "DE/", exptStr, "_zinbSNN_sconeNorm_pairsDE500.txt"))

plotHeatmap(log2(sconeNorm[unique(top5_zinb$IndexInOriginal),]+1), clusterSamplesData = ce@dendro_samples, clusterFeatures=T, sampleData=data.frame(clusters=z_cl, batch=colData(ce)[,1]), clusterLegend=list(clusters=colRKC, batch=colb), main=paste("oneVall5_SNNzinb1Kvar_sconeNorm"))
pdf(file="output/viz/heatmap_oneVall5_SNNzinb1Kvar_sconeNorm.pdf", height=8.5, width=11)
plotHeatmap(log2(sconeNorm[unique(top5_zinb$IndexInOriginal),]+1), clusterSamplesData = ce@dendro_samples, clusterFeatures=T, sampleData=data.frame(clusters=z_cl, batch=colData(filtered)[,1]), clusterLegend=list(clusters=colRKC, batch=colb), main=paste("oneVall5_SNNzinb1Kvar_sconeNorm")) 
dev.off()



######05/09/18
z_cl <- as.factor(as.numeric(z_cl))
names(z_cl) <- colnames(sconeNorm)
zinbcl1 <- ClusterExperiment(sconeNorm, clusters=z_cl, transformation = log1p)
zinbcl1 <- makeDendrogram(zinbcl1, reduceMethod = "var", nDims=1000)
plotDendrogram(zinbcl1)


pdf(file=paste0(outDir, "viz/", exptStr, "_heatmap_oneVall10_SNNzinb_sconeNorm.pdf"), height=8.5, width=11)
clusterExperiment::plotHeatmap(log2(sconeNorm[unique(top10_zinb$IndexInOriginal),]+1), clusterSamplesData = zinbcl1@dendro_samples, clusterFeatures=T, sampleData=data.frame(clusters=z_cl, batch=colData(filtered)[,1]), clusterLegend=list(clusters=colRKC, batch=colb), main=paste(exptStr, "_oneVall10_SNNzinb_sconeNorm")) 
dev.off()


top5pairs_zinb <- getBestFeatures(sconeNorm, cluster=z_cl, isCount=TRUE, contrastType = "Pairs", number=5)
pdf(file=paste0(outDir, "viz/", exptStr, "_heatmap_pairWise5_SNNzinb_sconeNorm.pdf"), height=8.5, width=11)
plotHeatmap(log2(sconeNorm[unique(top5pairs_zinb$IndexInOriginal),]+1), clusterSamplesData = zinbcl1@dendro_samples, clusterFeatures=T, sampleData=data.frame(clusters=z_cl, batch=colData(filtered)[,1]), clusterLegend=list(clusters=colRKC, batch=colb), main=paste("pairWise5_SNNzinb1Kvar_sconeNorm"))
dev.off()

plotHeatmap(log2(sconeNorm[unique(top5pairs_zinb$IndexInOriginal),]+1), clusterSamplesData = ce@dendro_samples, clusterFeatures=T, sampleData=data.frame(clusters=z_cl, batch=colData(ce)[,1]), clusterLegend=list(clusters=colRKC, batch=colb), main=paste("pairWise5_SNNzinb1Kvar_sconeNorm"))
pdf(file="output/viz/heatmap_pairWise5_SNNzinb1Kvar_sconeNorm.pdf", height=8.5, width=11)
plotHeatmap(log2(sconeNorm[unique(top5pairs_zinb$IndexInOriginal),]+1), clusterSamplesData = ce@dendro_samples, clusterFeatures=T, sampleData=data.frame(clusters=z_cl, batch=colData(filtered)[,1]), clusterLegend=list(clusters=colRKC, batch=colb), main=paste("pairWise5_SNNzinb1Kvar_sconeNorm")) 
dev.off()
###############################################
#######################

#####-----Using sconeNorm counts for all DE
#####-----DE by cluster for SNN clustering of sconeNorm 50 PCs

primaryClusterIndex(ce) <- 2
ce <- makeDendrogram(ce, dimReduce="PCA", ndims=10)
plotDendrogram(ce)

top5_scone50 <- getBestFeatures(sconeNorm, cluster=scone50_cl, isCount=TRUE, contrastType = "OneAgainstAll", number=5)
top1K_scone50 <- getBestFeatures(sconeNorm, cluster=scone50_cl, isCount=TRUE, contrastType = "OneAgainstAll", number=1000)
write.table(top1K_scone50, sep="\t", quote=FALSE,
            file = "output/DE/RCOB_scone50PCSNN_sconeNorm_oneVallDE1K.txt")

top5pairs_scone50 <- getBestFeatures(sconeNorm, cluster=scone50_cl, isCount=TRUE, contrastType = "Pairs", number=5)
top1Kpairs_scone50 <- getBestFeatures(sconeNorm, cluster=scone50_cl, isCount=TRUE, contrastType = "Pairs", number=1000)
write.table(top1Kpairs_scone50, sep="\t", quote=FALSE,
            file = "output/DE/RCOB_scone50PCSNN_sconeNorm_pairsDE1K.txt")

plotHeatmap(log2(sconeNorm[unique(top5_scone50$IndexInOriginal),]+1), clusterSamplesData = ce@dendro_samples, clusterFeatures=T, sampleData=data.frame(clusters=scone50_cl, batch=colData(ce)[,1]), clusterLegend=list(clusters=colRKC, batch=colb), main=paste("oneVall5_SNNscone50PC_sconeNorm"))
pdf(file="output/viz/heatmap_oneVall5_SNNscone50PC_sconeNorm.pdf", height=8.5, width=11)
plotHeatmap(log2(sconeNorm[unique(top5_scone50$IndexInOriginal),]+1), clusterSamplesData = ce@dendro_samples, clusterFeatures=T, sampleData=data.frame(clusters=scone50_cl, batch=colData(filtered)[,1]), clusterLegend=list(clusters=colRKC, batch=colb), main=paste("oneVall5_SNNscone50PC_sconeNorm")) 
dev.off()

plotHeatmap(log2(sconeNorm[unique(top5pairs_scone50$IndexInOriginal),]+1), clusterSamplesData = ce@dendro_samples, clusterFeatures=T, sampleData=data.frame(clusters=scone50_cl, batch=colData(ce)[,1]), clusterLegend=list(clusters=colRKC, batch=colb), main=paste("pairWise5_SNNscone50PC_sconeNorm"))
pdf(file="output/viz/heatmap_pairWise5_SNNscone50PC_sconeNorm.pdf", height=8.5, width=11)
plotHeatmap(log2(sconeNorm[unique(top5pairs_scone50$IndexInOriginal),]+1), clusterSamplesData = ce@dendro_samples, clusterFeatures=T, sampleData=data.frame(clusters=scone50_cl, batch=colData(filtered)[,1]), clusterLegend=list(clusters=colRKC, batch=colb), main=paste("pairWise5_SNNscone50PC_sconeNorm")) 
dev.off()
#####################
#####################
top50_scone50 <- getBestFeatures(sconeNorm, cluster=scone50_cl, isCount=TRUE, contrastType = "OneAgainstAll", number=50)
plotHeatmap(log2(sconeNorm[unique(top50_scone50$IndexInOriginal),]+1), clusterSamplesData = ce@dendro_samples, clusterFeatures=T, sampleData=data.frame(clusters=scone50_cl, batch=colData(ce)[,1]), clusterLegend=list(clusters=colRKC, batch=colb), main=paste("oneVall50_SNNscone50PC_sconeNorm"))
pdf(file="output/viz/heatmap_oneVall50_SNNscone50PC_sconeNorm.pdf", height=8.5, width=11)
plotHeatmap(log2(sconeNorm[unique(top50_scone50$IndexInOriginal),]+1), clusterSamplesData = ce@dendro_samples, clusterFeatures=T, sampleData=data.frame(clusters=scone50_cl, batch=colData(filtered)[,1]), clusterLegend=list(clusters=colRKC, batch=colb), main=paste("oneVall50_SNNscone50PC_sconeNorm")) 
dev.off()
top100_scone50 <- getBestFeatures(sconeNorm, cluster=scone50_cl, isCount=TRUE, contrastType = "OneAgainstAll", number=100)
pdf(file="output/viz/heatmap_oneVall100_SNNscone50PC_sconeNorm.pdf", height=8.5, width=11)
plotHeatmap(log2(sconeNorm[unique(top100_scone50$IndexInOriginal),]+1), clusterSamplesData = ce@dendro_samples, clusterFeatures=T, sampleData=data.frame(clusters=scone50_cl, batch=colData(filtered)[,1]), clusterLegend=list(clusters=colRKC, batch=colb), main=paste("oneVall100_SNNscone50PC_sconeNorm")) 
dev.off()

save(top1K_scone50, top1K_zinb, top1Kpairs_scone50, top1Kpairs_zinb, z_cl, scone50_cl, file="RCOB_DE_by_Cluster.Rda")
save(ce, z_cl, scone50_cl, filtCounts, W, pca_sconeNorm, sconeNorm, file="RCOB_ce.Rda")

```

###Visualizations based on cluster specific DE
####top 50 DE genes in putative tufted cell cluster
####Using the SNN clustering of sconeNorm 50PCs
#####tufted cell cluster is cluster 4
```{r zoomingIntoDEbyCluster}

primaryClusterIndex(ce) <- 2
ce <- makeDendrogram(ce, dimReduce="PCA", ndims=10)
plotDendrogram(ce)

tuftedDE <- top1K_scone50[top1K_scone50$Contrast=="Cl04-(Cl01+Cl02+Cl03+Cl05+Cl06+Cl07+Cl08+Cl09+Cl10)/9",]
tuftedDE <- tuftedDE[order(tuftedDE$logFC, decreasing=TRUE),]
#########---------###########
#############################----top 50 oneVall DE genes enriched in cluster 4 (tufted)
plotHeatmap(log2(sconeNorm[tuftedDE$Feature[1:50],]+1), clusterSamplesData = ce@dendro_samples, clusterFeatures=T, sampleData=data.frame(clusters=scone50_cl, batch=colData(ce)[,1]), clusterLegend=list(clusters=colRKC, batch=colb), main=paste("cluster4(tufted)DEtop50_SNNscone50PC_sconeNorm"))
pdf(file="output/viz/heatmap_cluster4(tufted)DEtop50_SNNscone50PC_sconeNorm.pdf", height=8.5, width=11)
plotHeatmap(log2(sconeNorm[tuftedDE$Feature[1:50],]+1), clusterSamplesData = ce@dendro_samples, clusterFeatures=T, sampleData=data.frame(clusters=scone50_cl, batch=colData(ce)[,1]), clusterLegend=list(clusters=colRKC, batch=colb), main=paste("cluster4(tufted)DEtop50_SNNscone50PC_sconeNorm"))
dev.off()
#########---------###########
#############################----top 100 oneVall DE genes enriched in cluster 4 (tufted)
plotHeatmap(log2(sconeNorm[tuftedDE$Feature[1:100],]+1), clusterSamplesData = ce@dendro_samples, clusterFeatures=T, sampleData=data.frame(clusters=scone50_cl, batch=colData(ce)[,1]), clusterLegend=list(clusters=colRKC, batch=colb), main=paste("cluster4(tufted)DEtop100_SNNscone50PC_sconeNorm"))
pdf(file="output/viz/heatmap_cluster4(tufted)DEtop100_SNNscone50PC_sconeNorm.pdf", height=8.5, width=11)
plotHeatmap(log2(sconeNorm[tuftedDE$Feature[1:100],]+1), clusterSamplesData = ce@dendro_samples, clusterFeatures=T, sampleData=data.frame(clusters=scone50_cl, batch=colData(ce)[,1]), clusterLegend=list(clusters=colRKC, batch=colb), main=paste("cluster4(tufted)DEtop100_SNNscone50PC_sconeNorm"))
dev.off()

############################
###########---------specific cluster enriched genes
Cl1DE <- top1K_scone50[top1K_scone50$Contrast=="Cl01-(Cl02+Cl03+Cl04+Cl05+Cl06+Cl07+Cl08+Cl09+Cl10)/9",]
Cl1DE <- Cl1DE[order(Cl1DE$logFC, decreasing=TRUE),]
Cl2DE <- top1K_scone50[top1K_scone50$Contrast=="Cl02-(Cl01+Cl03+Cl04+Cl05+Cl06+Cl07+Cl08+Cl09+Cl10)/9",]
Cl2DE <- Cl2DE[order(Cl2DE$logFC, decreasing=TRUE),]
Cl3DE <- top1K_scone50[top1K_scone50$Contrast=="Cl03-(Cl01+Cl02+Cl04+Cl05+Cl06+Cl07+Cl08+Cl09+Cl10)/9",]
Cl3DE <- Cl3DE[order(Cl3DE$logFC, decreasing=TRUE),]
Cl5DE <- top1K_scone50[top1K_scone50$Contrast=="Cl05-(Cl01+Cl02+Cl03+Cl04+Cl06+Cl07+Cl08+Cl09+Cl10)/9",]
Cl5DE <- Cl5DE[order(Cl5DE$logFC, decreasing=TRUE),]
Cl6DE <- top1K_scone50[top1K_scone50$Contrast=="Cl06-(Cl01+Cl02+Cl03+Cl04+Cl05+Cl07+Cl08+Cl09+Cl10)/9",]
Cl6DE <- Cl6DE[order(Cl6DE$logFC, decreasing=TRUE),]
Cl7DE <- top1K_scone50[top1K_scone50$Contrast=="Cl07-(Cl01+Cl02+Cl03+Cl04+Cl05+Cl06+Cl08+Cl09+Cl10)/9",]
Cl7DE <- Cl7DE[order(Cl7DE$logFC, decreasing=TRUE),]
Cl8DE <- top1K_scone50[top1K_scone50$Contrast=="Cl08-(Cl01+Cl02+Cl03+Cl04+Cl05+Cl06+Cl07+Cl09+Cl10)/9",]
Cl8DE <- Cl8DE[order(Cl8DE$logFC, decreasing=TRUE),]
Cl9DE <- top1K_scone50[top1K_scone50$Contrast=="Cl09-(Cl01+Cl02+Cl03+Cl04+Cl05+Cl06+Cl07+Cl08+Cl10)/9",]
Cl9DE <- Cl9DE[order(Cl9DE$logFC, decreasing=TRUE),]
Cl10DE <- top1K_scone50[top1K_scone50$Contrast=="Cl10-(Cl01+Cl02+Cl03+Cl04+Cl05+Cl06+Cl07+Cl08+Cl09)/9",]
Cl10DE <- Cl10DE[order(Cl10DE$logFC, decreasing=TRUE),]

plotHeatmap(log2(sconeNorm[Cl1DE$Feature[1:100],]+1), clusterSamplesData = ce@dendro_samples, clusterFeatures=T, sampleData=data.frame(clusters=scone50_cl, batch=colData(ce)[,1]), clusterLegend=list(clusters=colRKC, batch=colb), main=paste("cluster1_DEtop100_SNNscone50PC_sconeNorm"))
pdf(file="output/viz/heatmap_cluster1_DEtop100_SNNscone50PC_sconeNorm.pdf", height=8.5, width=11)
plotHeatmap(log2(sconeNorm[Cl1DE$Feature[1:100],]+1), clusterSamplesData = ce@dendro_samples, clusterFeatures=T, sampleData=data.frame(clusters=scone50_cl, batch=colData(ce)[,1]), clusterLegend=list(clusters=colRKC, batch=colb), main=paste("cluster1_DEtop100_SNNscone50PC_sconeNorm"))
dev.off()
plotHeatmap(log2(sconeNorm[Cl2DE$Feature[1:100],]+1), clusterSamplesData = ce@dendro_samples, clusterFeatures=T, sampleData=data.frame(clusters=scone50_cl, batch=colData(ce)[,1]), clusterLegend=list(clusters=colRKC, batch=colb), main=paste("cluster2_DEtop100_SNNscone50PC_sconeNorm"))
pdf(file="output/viz/heatmap_cluster2_DEtop100_SNNscone50PC_sconeNorm.pdf", height=8.5, width=11)
plotHeatmap(log2(sconeNorm[Cl2DE$Feature[1:100],]+1), clusterSamplesData = ce@dendro_samples, clusterFeatures=T, sampleData=data.frame(clusters=scone50_cl, batch=colData(ce)[,1]), clusterLegend=list(clusters=colRKC, batch=colb), main=paste("cluster2_DEtop100_SNNscone50PC_sconeNorm"))
dev.off()
plotHeatmap(log2(sconeNorm[Cl3DE$Feature[1:100],]+1), clusterSamplesData = ce@dendro_samples, clusterFeatures=T, sampleData=data.frame(clusters=scone50_cl, batch=colData(ce)[,1]), clusterLegend=list(clusters=colRKC, batch=colb), main=paste("cluster3_DEtop100_SNNscone50PC_sconeNorm"))
pdf(file="output/viz/heatmap_cluster3_DEtop100_SNNscone50PC_sconeNorm.pdf", height=8.5, width=11)
plotHeatmap(log2(sconeNorm[Cl3DE$Feature[1:100],]+1), clusterSamplesData = ce@dendro_samples, clusterFeatures=T, sampleData=data.frame(clusters=scone50_cl, batch=colData(ce)[,1]), clusterLegend=list(clusters=colRKC, batch=colb), main=paste("cluster3_DEtop100_SNNscone50PC_sconeNorm"))
dev.off()
plotHeatmap(log2(sconeNorm[Cl5DE$Feature[1:100],]+1), clusterSamplesData = ce@dendro_samples, clusterFeatures=T, sampleData=data.frame(clusters=scone50_cl, batch=colData(ce)[,1]), clusterLegend=list(clusters=colRKC, batch=colb), main=paste("cluster5_DEtop100_SNNscone50PC_sconeNorm"))
pdf(file="output/viz/heatmap_cluster5_DEtop100_SNNscone50PC_sconeNorm.pdf", height=8.5, width=11)
plotHeatmap(log2(sconeNorm[Cl5DE$Feature[1:100],]+1), clusterSamplesData = ce@dendro_samples, clusterFeatures=T, sampleData=data.frame(clusters=scone50_cl, batch=colData(ce)[,1]), clusterLegend=list(clusters=colRKC, batch=colb), main=paste("cluster5_DEtop100_SNNscone50PC_sconeNorm"))
dev.off()
plotHeatmap(log2(sconeNorm[Cl6DE$Feature[1:100],]+1), clusterSamplesData = ce@dendro_samples, clusterFeatures=T, sampleData=data.frame(clusters=scone50_cl, batch=colData(ce)[,1]), clusterLegend=list(clusters=colRKC, batch=colb), main=paste("cluster6_DEtop100_SNNscone50PC_sconeNorm"))
pdf(file="output/viz/heatmap_cluster6_DEtop100_SNNscone50PC_sconeNorm.pdf", height=8.5, width=11)
plotHeatmap(log2(sconeNorm[Cl6DE$Feature[1:100],]+1), clusterSamplesData = ce@dendro_samples, clusterFeatures=T, sampleData=data.frame(clusters=scone50_cl, batch=colData(ce)[,1]), clusterLegend=list(clusters=colRKC, batch=colb), main=paste("cluster6_DEtop100_SNNscone50PC_sconeNorm"))
dev.off()
plotHeatmap(log2(sconeNorm[Cl7DE$Feature[1:100],]+1), clusterSamplesData = ce@dendro_samples, clusterFeatures=T, sampleData=data.frame(clusters=scone50_cl, batch=colData(ce)[,1]), clusterLegend=list(clusters=colRKC, batch=colb), main=paste("cluster7_DEtop100_SNNscone50PC_sconeNorm"))
pdf(file="output/viz/heatmap_cluster7_DEtop100_SNNscone50PC_sconeNorm.pdf", height=8.5, width=11)
plotHeatmap(log2(sconeNorm[Cl7DE$Feature[1:100],]+1), clusterSamplesData = ce@dendro_samples, clusterFeatures=T, sampleData=data.frame(clusters=scone50_cl, batch=colData(ce)[,1]), clusterLegend=list(clusters=colRKC, batch=colb), main=paste("cluster7_DEtop100_SNNscone50PC_sconeNorm"))
dev.off()
plotHeatmap(log2(sconeNorm[Cl8DE$Feature[1:100],]+1), clusterSamplesData = ce@dendro_samples, clusterFeatures=T, sampleData=data.frame(clusters=scone50_cl, batch=colData(ce)[,1]), clusterLegend=list(clusters=colRKC, batch=colb), main=paste("cluster8_DEtop100_SNNscone50PC_sconeNorm"))
pdf(file="output/viz/heatmap_cluster8_DEtop100_SNNscone50PC_sconeNorm.pdf", height=8.5, width=11)
plotHeatmap(log2(sconeNorm[Cl8DE$Feature[1:100],]+1), clusterSamplesData = ce@dendro_samples, clusterFeatures=T, sampleData=data.frame(clusters=scone50_cl, batch=colData(ce)[,1]), clusterLegend=list(clusters=colRKC, batch=colb), main=paste("cluster8_DEtop100_SNNscone50PC_sconeNorm"))
dev.off()
plotHeatmap(log2(sconeNorm[Cl9DE$Feature[1:100],]+1), clusterSamplesData = ce@dendro_samples, clusterFeatures=T, sampleData=data.frame(clusters=scone50_cl, batch=colData(ce)[,1]), clusterLegend=list(clusters=colRKC, batch=colb), main=paste("cluster9_DEtop100_SNNscone50PC_sconeNorm"))
pdf(file="output/viz/heatmap_cluster9_DEtop100_SNNscone50PC_sconeNorm.pdf", height=8.5, width=11)
plotHeatmap(log2(sconeNorm[Cl9DE$Feature[1:100],]+1), clusterSamplesData = ce@dendro_samples, clusterFeatures=T, sampleData=data.frame(clusters=scone50_cl, batch=colData(ce)[,1]), clusterLegend=list(clusters=colRKC, batch=colb), main=paste("cluster9_DEtop100_SNNscone50PC_sconeNorm"))
dev.off()
plotHeatmap(log2(sconeNorm[Cl10DE$Feature[1:100],]+1), clusterSamplesData = ce@dendro_samples, clusterFeatures=T, sampleData=data.frame(clusters=scone50_cl, batch=colData(ce)[,1]), clusterLegend=list(clusters=colRKC, batch=colb), main=paste("cluster10_DEtop100_SNNscone50PC_sconeNorm"))
pdf(file="output/viz/heatmap_cluster10_DEtop100_SNNscone50PC_sconeNorm.pdf", height=8.5, width=11)
plotHeatmap(log2(sconeNorm[Cl10DE$Feature[1:100],]+1), clusterSamplesData = ce@dendro_samples, clusterFeatures=T, sampleData=data.frame(clusters=scone50_cl, batch=colData(ce)[,1]), clusterLegend=list(clusters=colRKC, batch=colb), main=paste("cluster10_DEtop100_SNNscone50PC_sconeNorm"))
dev.off()

```

###Simple Conclusions: 1. cluster 4 appears to be tufted cells and there appear to be several tufted specific genes, some strongly expressed, some less so; 2. it appears that cluster 2 has some cell cycle genes?????? (Cdkn1c, Ccnd1); 3. There appear to be several different types of projection neurons based on clustering and DE. 

###Analysis of RSEC clustering
```{r RSEC_scone50PC}
#####-----First, looking at scone 50PC based RSEC clustering; cmp for clusterManyPCA
load("dataObjects/RCOB_sconeClusterMany.Rda")

cmp <- combineMany(cl, proportion = 0.5, clusterFunction = "hierarchical01", minSize = 5, clusterLabel="combineMany,0.5")

plotClusters(cmp)

colData(cmp) <- colData(ce)
plotClusters(cmp, sampleData = "batch", colPalette=c(bigPalette, rainbow(50)))
cmp <- makeDendrogram(cmp, dimReduce="PCA", ndims=10)
plotDendrogram(cmp)

mergeClusters(cmp,mergeMethod="adjP",plotInfo="mergeMethod")
cmp <- mergeClusters(cmp, mergeMethod = "adjP", cutoff = 0.04, plotInfo="mergeMethod")
cmp <- makeDendrogram(cmp, dimReduce="PCA", ndims=10)
plotDendrogram(cmp)
final_hier_sc <- factor(clusterMatrixNamed(cmp)[,"mergeClusters"],
                      levels=clusterLegend(cmp)[["mergeClusters"]][, "name"])
names(final_hier_sc) <- colnames(cmp)
write.table(final_hier_sc, file="output/clustering/RSEC_merged_adjP04_hier_clusters.txt", quote=FALSE, sep='\t',
            col.names = FALSE)

pdf(file="output/RSECsc50mergeHeatmap.pdf")
plotHeatmap(cmp,main="Heatmap with clusterMany")
dev.off()

markers <- c("Thy1", "Gad1", "Gad2", "Fxyd6", "Fxyd7", "Bcl2", "Pcp4", "Meis2", "Kcna2", "Gabra1", "Grin2b", "Dnajc2", "Vps8", "Calb2", "Fstl5", "Camk2n1", "Gls", "Reln", "Slc17a7", "Fst", "Coch", "tdTomato", "CreER", "Kcnab3", "Ctgf", "Ly6g6e", "Pip5k1b", "Nxph4", "Cbln2", "Cadps2", "Id2", "Samsn1", "Mgst3", "Coro1a", "Shisa3", "Igfbp5", "Cbln4", "Ntng1", "Coro6", "Nrp2", "Gabra2", "Doc2b")
pdf(file="output/RSECmergeAdjP05_sc50heatmapCMP_markers.pdf")
plotHeatmap(log2(assay(cmp)[markers,]+1), clusterSamplesData = cmp@dendro_samples, clusterFeatures=T, sampleData=data.frame(RSECsc50PC_Cl=final_hier_sc, SNNsc50PC_Cl=scone50_cl, SNNzinbCl=z_cl, batch=colData(ce)[,1]), clusterLegend=list(RSECsc50PC_Cl=colRKC, SNNsc50PC_Cl=colRKC, SNNzinbCl=colRKC, batch=colb), labCol=NA,  unassignedColor = "white", main=paste("RSEC scone50PCs, select markers"))
dev.off()

hwh_rm <- which(final_hier_sc=="-1")
pca <- prcomp(t(logSconeNorm[,-hwh_rm]))

plot(pca$x, pch=19, col=colRKC[final_hier_sc[-hwh_rm]], main="RSEC_merged_scone50PCs, PCA, hierarchical01")
legend("topleft", setdiff(levels(final_hier_sc),'-1'), fill=colRKC, cex = 0.5)

vars <- rowVars(logSconeNorm)
names(vars) <- rownames(logSconeNorm)
vars <- sort(vars, decreasing = TRUE)

pdf(file="output/viz/tSNE_RSEC_mergeAdjP05_scone50PC.pdf")
tsne_1000 <- Rtsne(t(logSconeNorm[names(vars)[1:1000],-hwh_rm]),
                   max_iter=5000, perplexity = 30)
plot(tsne_1000$Y, pch=19, cex = 0.4, col=colRKC[final_hier_sc[-hwh_rm]], main="t-SNE, RSEC scone50PC, 1000 most variable genes, hier01")
legend("topright", setdiff(levels(final_hier_sc),'-1'), fill=colRKC, cex = 0.5)

tsne_500 <- Rtsne(t(logSconeNorm[names(vars)[1:500],-hwh_rm]),
                   max_iter=5000, perplexity = 30)
plot(tsne_500$Y, pch=19, cex = 0.4, col=colRKC[final_hier_sc[-hwh_rm]], main="t-SNE, RSEC scone50PC, 500 most variable genes, hier01")
legend("topright", setdiff(levels(final_hier_sc),'-1'), fill=colRKC, cex = 0.5)

tsne_250 <- Rtsne(t(logSconeNorm[names(vars)[1:250],-hwh_rm]),
                   max_iter=5000, perplexity = 30)
plot(tsne_250$Y, pch=19, cex = 0.4, col=colRKC[final_hier_sc[-hwh_rm]], main="t-SNE, RSEC scone50PC, 250 most variable genes, hier01")
legend("topright", setdiff(levels(final_hier_sc),'-1'), fill=colRKC, cex = 0.5)

tsne_rsecScone50PC50 <- Rtsne(pca$x[,1:50], pca=FALSE, max_iter = 5000)
plot(tsne_rsecScone50PC50$Y, pch=19, cex = 0.4, col=colRKC[final_hier_sc[-hwh_rm]], main="t-SNE, RSEC scone50PC, 50 PCs tSNE, hier01")
legend("topright", setdiff(levels(final_hier_sc),'-1'), fill=colRKC, cex = 0.5)

# tsne_rsecScone50PC20 <- Rtsne(pca$x[,1:20], pca=FALSE, max_iter = 5000)
# plot(tsne_rsecScone50PC20$Y, pch=19, cex = 0.4, col=pal[final_hier_sc[-hwh_rm]], main="t-SNE, RSEC scone50PC, 20 PCs tSNE, hier01")
dev.off()

```

###Analysis of RSEC clustering
```{r RSEC_zinb}
#####-----First, looking at scone 50PC based RSEC clustering; cmp for clusterManyPCA
load("dataObjects/RCOB_zinbClusterMany.Rda")

cmpz <- combineMany(cl, proportion = 0.5, clusterFunction = "hierarchical01", minSize = 5, clusterLabel="combineMany,0.5")

plotClusters(cmpz)

colData(cmpz) <- colData(ce)
plotClusters(cmpz, sampleData = "batch", colPalette=c(bigPalette, rainbow(50)))
cmpz <- makeDendrogram(cmpz, dimReduce="PCA", ndims=10)
plotDendrogram(cmpz)

mergeClusters(cmpz,mergeMethod="adjP",plotInfo="mergeMethod")
cmpz <- mergeClusters(cmpz, mergeMethod = "adjP", cutoff = 0.06, plotInfo="mergeMethod")
cmpz <- makeDendrogram(cmpz, dimReduce="PCA", ndims=10)
plotDendrogram(cmpz)
final_hier_sc <- factor(clusterMatrixNamed(cmpz)[,"mergeClusters"],
                      levels=clusterLegend(cmpz)[["mergeClusters"]][, "name"])
names(final_hier_zb) <- colnames(cmpz)
write.table(final_hier_zb, file="output/RSEC_merged_hier_clusters.txt", quote=FALSE, sep='\t',
            col.names = FALSE)

pdf(file="output/RSECzinbMergeHeatmap.pdf")
plotHeatmap(cmpz,main="Heatmap with clusterMany")
dev.off()

markers <- c("Thy1", "Gad1", "Gad2", "Fxyd6", "Fxyd7", "Bcl2", "Pcp4", "Meis2", "Kcna2", "Gabra1", "Grin2b", "Dnajc2", "Vps8", "Calb2", "Fstl5", "Camk2n1", "Gls", "Reln", "Slc17a7", "Fst", "Coch", "tdTomato", "CreER", "Kcnab3", "Ctgf", "Ly6g6e", "Pip5k1b", "Nxph4", "Cbln2", "Cadps2", "Id2", "Samsn1", "Mgst3", "Coro1a", "Shisa3", "Igfbp5", "Cbln4", "Ntng1", "Coro6", "Nrp2", "Gabra2", "Doc2b")
pdf(file="output/RSECzbheatmapCMPZ_markers.pdf")
plotHeatmap(log2(assay(cmpz)[markers,]+1), clusterSamplesData = cmp@dendro_samples, clusterFeatures=T, sampleData=data.frame(RSECzinb_Cl=final_hier_sc, SNNzinb_Cl=scone50_cl, SNNzinbCl=z_cl, batch=colData(ce)[,1]), clusterLegend=list(RSECzb_Cl=colRKC, SNNsc50PC_Cl=colRKC, SNNzinbCl=colRKC, batch=colb), main=paste("RSEC scone50PCs, select markers"))
dev.off()

zwh_rm <- which(final_hier_sc=="-1")
pca <- prcomp(t(logSconeNorm[,-zwh_rm]))

plot(W, pch=19, col=colRKC[final_hier_sc[-zwh_rm]], main="RSEC_merged_scone50PCs, PCA, hierarchical01")
legend("topleft", setdiff(levels(final_hier_zb),'-1'), fill=colRKC, cex = 0.5)

vars <- rowVars(logSconeNorm)
names(vars) <- rownames(logSconeNorm)
vars <- sort(vars, decreasing = TRUE)

tsne_1000zb <- Rtsne(t(logSconeNorm[names(vars)[1:1000],-zwh_rm]),
                   max_iter=5000, perplexity = 30)
plot(tsne_1000zb$Y, pch=19, cex = 0.4, col=colRKC[final_hier_zb[-zwh_rm]], main="t-SNE, RSEC scone50PC, 1000 most variable genes, hier01")

tsne_500zb <- Rtsne(t(logSconeNorm[names(vars)[1:500],-zwh_rm]),
                   max_iter=5000, perplexity = 30)
plot(tsne_500zb$Y, pch=19, cex = 0.4, col=colRKC[final_hier_zb[-zwh_rm]], main="t-SNE, RSEC scone50PC, 500 most variable genes, hier01")

tsne_250zb <- Rtsne(t(logSconeNorm[names(vars)[1:250],-zwh_rm]),
                   max_iter=5000, perplexity = 30)
plot(tsne_250zb$Y, pch=19, cex = 0.4, col=colRKC[final_hier_zb[-zwh_rm]], main="t-SNE, RSEC scone50PC, 250 most variable genes, hier01")

tsne_rsecZinb_W10 <- Rtsne(W[,1:10], pca=FALSE, max_iter = 5000)
plot(tsne_rsecScone50PC50$Y, pch=19, cex = 0.4, col=colRKC[final_hier_zb[-zwh_rm]], main="t-SNE, RSEC zinb,  W10_tSNE, hier01")


```


